\section{Live Hugging Face rerun (v2.1, \texttt{core4} scope)}\label{live-hf-rerun-v2.1}

The v2.0 battery (Sections \ref{results-layer-1-public-benchmarks-h1h7}--\ref{summary-of-l1} and
Appendix \ref{reproducibility-manifest}) runs the \emph{synthetic-fallback}
path of every L1 adapter by default so the whole validation re-runs on a
laptop in under three minutes. A reviewer can reasonably ask: ``does the
paired delta hold when the adapter is forced to actually pull data from
Hugging Face at the current commit of each upstream dataset?'' This
appendix documents a \textbf{preregistered, strictly live} rerun of the
four headline L1 bundles against live Hugging Face data, run between
2026-04-20 and 2026-04-21, and reports exactly what that pass found.

\subsection{Pre-registration and strictness}\label{live-hf-prereg}

The run is locked by a pre-registration in
\passthrough{\lstinline!experiments/v2/p6a/specs_live.py!} that fixes, for
every bundle in the \passthrough{\lstinline!core4!} scope:

\begin{itemize}
\tightlist
\item \(N=15\) examples per (model, seed, tier) cell.
\item Two seeds: \passthrough{\lstinline!(0, 1)!}.
\item Two models: \passthrough{\lstinline!claude-haiku-4-5!} and
  \passthrough{\lstinline!claude-sonnet-4-6!}.
\item Two RULER tiers: \passthrough{\lstinline!(8192, 16384)!} tokens.
  (The upstream dataset
  \passthrough{\lstinline!simonjegou/ruler!} publishes only
  \(\{4096, 8192, 16384\}\); the synthetic-fallback path still supports
  32\,K, but the live dataset does not.)
\item \passthrough{\lstinline!strict\_hf=True!} on every adapter, which
  converts any Hugging Face loading failure into a
  \passthrough{\lstinline!ProvenanceIntegrityError!} rather than a
  silent fall-through to synthetic data.
\end{itemize}

The CLI runner
\passthrough{\lstinline!experiments/v2/p6a/run\_live\_hf.py!} enforces the
pre-registration by refusing to start if any of $N$, seeds, models, or
tiers were overridden on the command line under
\passthrough{\lstinline!--scope core4!} or
\passthrough{\lstinline!--scope full\_battery!}. It also executes a
``probe-then-verify'' step: it loads one example per bundle before any
API calls go out, reads back
\passthrough{\lstinline!provenance.source!}, and aborts with exit code
\texttt{3} if that field is not the literal string
\passthrough{\lstinline!"huggingface"!}.

\subsection{Provenance}\label{live-hf-provenance}

Each bundle's JSON output carries a
\passthrough{\lstinline!provenance!} block pinning both the Hugging Face
dataset id and the \emph{commit SHA} of that dataset at run time (the
SHAs below were resolved through the Hugging Face Hub API at
consolidation time and stamped into
\passthrough{\lstinline!\_summary\_live.json!}):

\begin{itemize}
\tightlist
\item \passthrough{\lstinline!simonjegou/ruler!} @
  \texttt{24adceac8a0e6532936e8d721cd9e9084d2e4686}, split
  \texttt{test}, task filter \texttt{niah\_single\_1}.
\item \passthrough{\lstinline!truthfulqa/truthful\_qa!} @
  \texttt{741b8276f2d1982aa3d5b832d3ee81ed3b896490}, config
  \texttt{generation}.
\item \passthrough{\lstinline!princeton-nlp/SWE-bench\_Verified!} @
  \texttt{c104f840cc67f8b6eec6f759ebc8b2693d585d4a}.
\end{itemize}

Together with the harness \passthrough{\lstinline!git\_sha!} (also
stamped on every record) this fixes every input bit that enters any
statistic in Table\,\ref{tab:t8_live_hf}.

\subsection{Scope: why \texttt{core4} and not
\texttt{full\_battery}}\label{live-hf-scope}

We had planned a 7-bundle \passthrough{\lstinline!full\_battery!} run.
After the first Anthropic 5-hour rolling window exhausted without
finishing SWE-bench, we \textbf{amended the pre-registration
pre-execution} to the \textbf{minimum four} bundles that are
\emph{individually} responsible for distinct headline claims in the
paper, so that a single rolling window's worth of budget would suffice
to complete the comparison at the same $N$, seeds, models, and tiers as
the full battery. The amendment is a \emph{protocol deviation}, not a
post-hoc selection on observed effect sizes: the four retained bundles
were chosen by \emph{claim coverage} (one bundle per independent L1
claim), the three dropped bundles are listed explicitly below, the
dropped-bundle checkpoints are archived rather than deleted, and the
$N$/seeds/models/tiers of the retained bundles are \emph{identical} to
the original 7-bundle pre-registration (the locking is still enforced at
the CLI, see §\ref{live-hf-prereg}). The amendment was made before any
of the four retained bundles completed; none of the numbers in
Table~\ref{tab:t8_live_hf} were available at the time of the narrowing.
The four retained bundles are:

\begin{itemize}
\tightlist
\item \passthrough{\lstinline!H1\_ruler\_8192\_live!} and
  \passthrough{\lstinline!H1\_ruler\_16384\_live!} --- \emph{one} tier
  alone suffices to assert the recall effect; \emph{two} tiers are
  needed to assert the Lost-in-the-Middle \emph{slope} claim.
\item \passthrough{\lstinline!H\_TQA\_live\_v2!} --- TruthfulQA, the
  hallucination / calibration claim.
\item \passthrough{\lstinline!H\_SWEB\_live\_n15!} --- SWE-bench
  Verified, the coding / agentic claim.
\end{itemize}

The three dropped bundles
(\passthrough{\lstinline!H1b\_ruler\_multi\_8192\_live!},
\passthrough{\lstinline!H1b\_ruler\_multi\_16384\_live!},
\passthrough{\lstinline!H\_HEQA\_live\_v2!}) are structural variants of
retained bundles and their checkpoints have been archived to
\passthrough{\lstinline!.cache/archive/!} rather than deleted.

\subsection{Results}\label{live-hf-results}

Headline numbers for all four bundles are in Table~\ref{tab:t8_live_hf}
(Section~\ref{live-hf-rerun-summary}); we discuss each row in turn here.

\textbf{H1 RULER, 8\,K tokens.} Treatment 1.0000, baseline 0.7667,
paired $\Delta=+0.2333$ (paired-permutation $p=0.0005$, $d_z=0.547$),
$N=60$. The 8\,K RULER effect \emph{reproduces on real
simonjegou/ruler data} at an effect size and paired $p$-value that
easily clear the pre-registered $d \ge 0.5$ / $p \le 0.05$ target gate
at \(N=60\) paired observations (15 instances $\times$ 2 seeds
$\times$ 2 models).

\textbf{H1 RULER, 16\,K tokens.} Treatment 1.0000, baseline 0.9333,
paired $\Delta=+0.0667$ ($p=0.1174$, $d_z=0.265$), $N=60$. The effect is
in the predicted direction but does \emph{not} clear the target gate at
$N=60$; the underlying distribution is highly skewed (baseline
already scores $0.9333$ on the easier live RULER 16\,K examples). A
larger $N$ would be required to assert the effect at 16\,K on this
revision of the dataset; we therefore do not claim a live 16\,K effect
and note the direction for the reader. The synthetic-fallback
v2.0 result at 32\,K in §\ref{h1-long-context-recall-ruler} used a
different tier and a different generator; it is not invalidated by this
row, but the paper's Lost-in-the-Middle \emph{slope} claim now rests
exclusively on the v2.0 synthetic-fallback evidence.

\textbf{H\_TQA\_live\_v2.} Treatment $0.0500$, baseline $0.0833$,
$\Delta=-0.0333$ ($p=0.7386$, $d_z=-0.091$), $N=60$. The live
TruthfulQA rerun is a \emph{null effect}: both arms score near-floor on
this revision at $N=60$, well below the gate. We report the null
honestly. The synthetic-fallback hallucination-class evidence in
§8 is not invalidated by this row --- different benchmark surface,
different adapter --- but the paper's calibration claim on TruthfulQA
specifically should be read as ``we could not reproduce an effect on
the live pull at $N=60$''.

\textbf{H\_SWEB\_live\_n15, haiku-only.} The full SWE-bench bundle is
$15 \times 2 \times 2 \times 2 = 120$ records. The
\passthrough{\lstinline!claude-haiku-4-5!} side completed in full
($60/60$, both seeds, both conditions). The
\passthrough{\lstinline!claude-sonnet-4-6!} side was then blocked at the
CLI layer --- see §\ref{live-hf-sweb-cli-limitation} below --- and we
report only the haiku slice here.

On 30 paired (seed, instance) observations for haiku, with CLI-aborted
attempts scored as $0$ on both sides per pre-registration: harness\_on
mean $0.1259$, harness\_off mean $0.0167$, paired $\Delta=+0.1093$
(paired-permutation $p=0.0322$, $d_z=0.488$). The $p$-value clears the
preregistered $p \le 0.05$ gate; the effect size sits
\emph{just below} the preregistered $d \ge 0.5$ gate ($d_z=0.488$) --- we
therefore treat SWE-bench-haiku as \emph{directionally} reproducing at
conventional significance rather than clearing the combined gate.

The direction matches the v2.0 deterministic
\passthrough{\lstinline!PatchSimulator!} result
(Section~\ref{headline-result} in
Section~\ref{results-layer-3-swe-bench-verified-ab-head-to-head-p6-c}) at
a smaller effect size, consistent with the
Section~\ref{discussion-limitations-and-threats-to-validity} prediction
that a real-LLM coder will compress the patch-simulator's gain.

\emph{Imputation sensitivity.} The $p=0.0322$ above is computed under
the pre-registered rule that CLI-aborted attempts score $0$ on
\emph{both} sides. Restricting to
\passthrough{\lstinline!clean\_pairs\_only!} --- (seed, instance) combos
where neither condition aborted --- leaves only
\textbf{1} clean paired observation (see
\passthrough{\lstinline!swe\_bench\_outcomes.json!}, field
\passthrough{\lstinline!clean\_pairs\_only!}), so essentially all of
the paired signal above comes from zero-imputed rows rather than from
observed harness-on / harness-off score pairs. The reader should
therefore treat the haiku SWE-bench row as \emph{suggestive under the
pre-registered imputation rule}, not as a large-$N$ reproduction.

\subsection{SWE-bench CLI-level limitation}\label{live-hf-sweb-cli-limitation}

Of 30 paired (seed, instance) attempts on the
\passthrough{\lstinline!claude-haiku-4-5!} side of
\passthrough{\lstinline!H\_SWEB\_live\_n15!}, \textbf{23 (77\%)}
aborted at the \passthrough{\lstinline!claude!} CLI layer under the
scheduler's 300\,s \passthrough{\lstinline!timeout\_s!} on \emph{both}
the \passthrough{\lstinline!harness\_on!} and
\passthrough{\lstinline!harness\_off!} conditions, at the same 23
(seed, instance) combos. On the
\passthrough{\lstinline!claude-sonnet-4-6!} side, 10/10 attempts
aborted before the second rolling window exhausted. In every aborted
case the CLI's NDJSON stream began with the
\passthrough{\lstinline!\{"type":"system","subtype":"hook\_started"\}!}
event (from the
\passthrough{\lstinline!SessionStart!} hook installed by the host
agent session) and then terminated without emitting the canonical
\passthrough{\lstinline!\{"type":"result"\}!} object that the scheduler
parses; the scheduler exhausts its 3-retry budget and logs the
\passthrough{\lstinline!SchedulerError!}.

The likely mechanism is that very large SWE-bench prompts
(\(\sim\)10\,--\,30\,K tokens of problem statement plus the
host-agent's own cached startup context) push the CLI past the
\passthrough{\lstinline!timeout\_s=300!} budget before the
SessionStart hook completes; the hook never unblocks, so the CLI
process ends without the control-plane \passthrough{\lstinline!result!}
event. Crucially this is a \emph{CLI-level infrastructure failure, not
a modeling result}: it does not indicate that the harness or the
baseline answered the problem incorrectly, only that the CLI never let
either of them answer. The failure is \emph{balanced} by construction
--- the same 23 (seed, instance) combos aborted on both conditions ---
so the pre-registered conservative scoring (errors = 0 on both sides)
does not introduce a directional bias, but it does squeeze the paired
signal into 7 (seed, instance) combos that actually ran one or both
sides. The reader should therefore treat the haiku-only row in
Table\,\ref{tab:t8_live_hf} as \emph{suggestive, not confirmatory} of
the v2.0 SWE-bench headline in §10.

We retain the record (a) because it still points in the predicted
direction ($\Delta=+0.1093$, $d_z=0.488$, paired $p=0.0322$) and (b)
because the exact failure mode is an artefact of the
claude CLI 2.1 plus a host-agent SessionStart hook --- both
independently recorded --- so a future rerun on a CLI release that
removes the hook-startup blocking, or a longer
\passthrough{\lstinline!timeout\_s!}, can close this gap without
changing any paper numbers downstream.

\subsection{Quota accounting}\label{live-hf-quota}

The scheduler wrote 342 CLI calls across the full rerun (three
accounts / three rolling windows), consuming 11.58\,M input tokens and
84\,566 output tokens on the last-reported cumulative window; the
\passthrough{\lstinline!next\_window\_at!} field in
\passthrough{\lstinline!\_summary\_live.json!} pins the Anthropic 5-hour
window that was active at the final consolidation time.

\subsection{Replay the live rerun}\label{live-hf-replay}

From a clean checkout at the same harness commit SHA, with
\passthrough{\lstinline!HF\_TOKEN!} set in \passthrough{\lstinline!.env!}:

\begin{lstlisting}[language=bash]
# 1. Verify the adapter-level strict-HF + probe behaviour.
uv run pytest -q tests/dev/ tests/test_benchmarks/ \
    tests/test_v2/test_p6a_callers.py tests/test_v2/test_p6a_run.py

# 2. Run the core4 live-HF battery with pre-registration enforced.
uv run python -m experiments.v2.p6a.run_live_hf \
    --scope core4 --live --continue-on-partial

# 3. Consolidate (idempotent; stamps hf_revision from the Hub API):
uv run python -m tools.dev.live_hf_consolidate

# 4. Compute the haiku-only SWE-bench side-car:
uv run python -m tools.dev.sweb_outcomes
\end{lstlisting}

Each step is idempotent; (2) resumes from the per-bundle JSONL
checkpoints under
\passthrough{\lstinline!.cache/live\_hf\_checkpoints/!} when interrupted
by an Anthropic rolling-window exhaustion. Expect a total wall-clock of
\(\sim\)60--90 minutes per completed 5-hour rolling window.

\subsection{v2.1.1 Power-extension addendum (pre-execution
amendment)}\label{live-hf-v2.1.1}

The v2.1 read left three conclusions contingent on the
\passthrough{\lstinline!N=15!} budget: RULER~16\,K
(\emph{directional but underpowered at $p=0.117$}), TruthfulQA
(\emph{null at $p=0.74$; CI wide enough to leave a small positive
effect compatible with the data}), and SWE-bench Verified
(\emph{partial at the CLI layer: 77\% of haiku attempts and 100\% of
sonnet attempts aborted inside the Claude CLI's
\passthrough{\lstinline!SessionStart!} hook under the scheduler's
300\,s subprocess timeout, not inside the model response}). We wrote a
\textbf{pre-execution amendment} --- filed under
\passthrough{\lstinline!specs\_live.py::power\_ext\_specs!} and
enforced at the CLI under \passthrough{\lstinline!--scope power\_ext!}
--- to address those three contingencies \emph{minimally} rather than
opening a new N:

\begin{itemize}
\tightlist
\item \textbf{\passthrough{\lstinline!H1\_ruler\_16384\_live\_ext!}}:
  $N$ doubled from $15$ to $30$ (per (model, seed, condition) cell),
  same two seeds, same two models, same strict-HF provenance. The
  RULER adapter's HF loader is deterministic in \passthrough{\lstinline!seed!}
  (standard-library \passthrough{\lstinline!random.Random(seed).shuffle!} over
  the full HF revision at the pinned commit), so the $N=30$ pull is a
  strict \emph{superset} of the $N=15$ pull at the same seed; the v2.1
  checkpoint JSONL is seeded verbatim into the \passthrough{\lstinline!\_ext!}
  file so the first 15 rows per cell replay as cache-hits and only
  the additional 15 bill.
\item \textbf{\passthrough{\lstinline!H\_TQA\_live\_v2\_ext!}}: $N$
  doubled from $15$ to $30$ with the same checkpoint-seeding trick.
  Goal: narrow the 95\% CI around zero to either confirm the null
  at $\ge 0.8$ power for $|d|\ge 0.30$ or surface a previously
  undetected effect.
\item \textbf{\passthrough{\lstinline!H\_SWEB\_live\_ext!}}: $N$
  held at the pre-registered $15$, seeds unchanged, models
  unchanged, scorer unchanged; the \emph{only} change is the CLI
  subprocess timeout --- raised from 300\,s to
  \passthrough{\lstinline!LIVE\_EXT\_SWEB\_TIMEOUT\_S = 900!}~s
  so the \passthrough{\lstinline!SessionStart!}-hook abort path
  from \S\ref{live-hf-sweb-cli-limitation} is eliminated. We
  explicitly refuse to extend $N$ on SWE-bench: \emph{even at a
  larger $N$ the heuristic scorer has a low per-instance ceiling,
  so a power extension would still be underpowered at reachable
  budgets, whereas the infrastructure fix converts the already-paid
  budget from CLI-aborts into scored patches}. The v2.1 JSONL is
  pre-filtered to drop error rows before being seeded into the
  \passthrough{\lstinline!\_ext!} checkpoint, so previously errored
  (model, seed, instance, condition) tuples re-bill under the new
  timeout.
\end{itemize}

The amendment was filed and the CLI pre-registration lock was
extended (see \passthrough{\lstinline!run\_live\_hf.py::\_enforce\_preregistration!}
with the \passthrough{\lstinline!power\_ext!} arm) \emph{before}
any power-extension row was billed. The three \passthrough{\lstinline!\_ext!}
bundles carry their own hypothesis ids (\passthrough{\lstinline!H1\_ext!},
\passthrough{\lstinline!H\_TQA\_v2\_ext!}, \passthrough{\lstinline!H\_SWEB\_ext!})
so downstream analysis can never confuse a v2.1 N=15 row with a
v2.1.1 N=30 row.

\subsubsection{What actually happened}

Anthropic's account-level rolling quota exhausted partway through the
amendment. After halting the run we treated the checkpoint JSONLs as
final and re-scored them offline (no further Claude CLI / Anthropic
API traffic; the consolidator still performs a cached read-only
\passthrough{\lstinline!HfApi().dataset\_info!} call per pinned
dataset to stamp \passthrough{\lstinline!provenance.hf\_revision!},
but this does not bill against Anthropic quota and is skipped when
the revision is already stamped on the record) with
\passthrough{\lstinline!tools/dev/score\_ext\_checkpoints.py!}. The
score payloads are written next to the \passthrough{\lstinline!\_partial.json!}
files as \passthrough{\lstinline!\_score.json!}; the consolidator
splices them into \passthrough{\lstinline!\_summary\_live.json!}.

\textbf{H1\_ruler\_16384\_live\_ext.} Pooled $n_\text{pair}=103$
(haiku fully extended to $n=60$; sonnet seed~0 extended to $n=28$
after two scheduler-level CLI aborts; sonnet seed~1 still at the
original $n=15$). Treatment mean $0.9903$, baseline $0.9417$,
paired $\Delta=+0.0485$, 95\% bootstrap CI \textbf{[+0.010,\,+0.087]
(excludes 0)}, paired-permutation $p=0.064$, Cohen's $d_z=0.225$.
Per-model: haiku $\Delta=+0.050$, $p=0.26$, $n=60$; sonnet
$\Delta=+0.047$, $p=0.51$, $n=43$. Neither preregistered gate
($\Delta\ge 0.05$ \emph{and} $p\le 0.05$) clears --- so the result
reads as \textbf{tighter interval, same verdict: directional but
underpowered}. The two sonnet seed-0 CLI aborts re-exposed the
\passthrough{\lstinline!SessionStart!}-hook mode from
\S\ref{live-hf-sweb-cli-limitation} on RULER prompts as well
(not only on SWE-bench); the v2.1.1 CLI-timeout fix was carried in
the \passthrough{\lstinline!power\_ext!} scope for RULER too, yet
the aborts still surfaced intermittently, suggesting the CLI hook
failure is partially non-timeout (a plain process exit on some
invocations). We include the two error rows in the paired totals
as scored 0 (per the pre-registered imputation convention); excluding
them moves pooled $p$ to $0.061$ and $d_z$ to $0.229$ --- same
verdict.

\textbf{H\_TQA\_live\_v2\_ext.} No new rows billed before the
quota halted. The checkpoint is byte-identical to the v2.1 N=15
TruthfulQA checkpoint; the re-score is the v2.1 result re-reported
at $n_\text{pair}=60$: $\Delta=-0.033$, $p=0.736$, $d_z=-0.091$.
The v2.1.1 read on TruthfulQA therefore adds \emph{no new
information}; the v2.1 null remains the operative verdict, and a
proper power-extension is left as pre-registered open work.

\textbf{H\_SWEB\_live\_ext.} The infra-rescue code ships
(\passthrough{\lstinline!--scheduler-timeout-s!} CLI flag wired
end-to-end, \passthrough{\lstinline!LIVE\_EXT\_SWEB\_TIMEOUT\_S!}
default to 900\,s, pre-registration lock extended) but only one
paired (seed, instance) observation billed before the quota halted
--- so the v2.1 haiku-only row in Table~\ref{tab:t8_live_hf}
remains the operative SWE-bench live read, and the new timeout
floor is shipped-but-unexercised. A future rerun on a refreshed
Anthropic window will exercise the fix without changing any
paper numbers upstream of SWE-bench.

\subsubsection{Pre-registration integrity and what the amendment
does not do}

The v2.1.1 amendment is a \emph{pre-execution} amendment: it was
filed and the CLI enforcement was wired \emph{before} any
\passthrough{\lstinline!\_ext!} row was billed, and the three
\passthrough{\lstinline!\_ext!} bundles are \emph{new labels},
not overrides of the v2.1 bundles. The v2.1 JSON outputs, the v2.1
\passthrough{\lstinline!SHA256SUMS!} lines, and Table~\ref{tab:t8_live_hf}
v2.1 rows remain byte-reproducible from the v2.1 commit tag. The
amendment also does not open SWE-bench to a larger $N$ (for the
reason above): the SWE-bench cell stays at the original
pre-registered $N=15$ / seeds $(0,1)$ / models / scorer and only
the subprocess timeout moves. We therefore treat the v2.1.1
numbers as \emph{companion reads} rather than as replacements
for the v2.1 rows, and surface both in Table~\ref{tab:t8_live_hf}.

\subsubsection{Replay the v2.1.1 run}

\begin{lstlisting}[language=bash]
# (1) optional: rebuild the seeded _ext checkpoints from the v2.1
# JSONLs (already committed; this step is idempotent):
cp .cache/live_hf_checkpoints/H1_ruler_16384_live.jsonl \
   .cache/live_hf_checkpoints/H1_ruler_16384_live_ext.jsonl
cp .cache/live_hf_checkpoints/H_TQA_live_v2.jsonl \
   .cache/live_hf_checkpoints/H_TQA_live_v2_ext.jsonl
# SWE seed file: drop rows with error != "":
python3 -c "
import json, pathlib
src = pathlib.Path('.cache/live_hf_checkpoints/H_SWEB_live_n15.jsonl')
dst = pathlib.Path('.cache/live_hf_checkpoints/H_SWEB_live_ext.jsonl')
keep = [l for l in src.read_text().splitlines()
        if l.strip() and not json.loads(l).get('error')]
dst.write_text('\n'.join(keep) + '\n')
"

# (2) run the power_ext scope (enforces pre-registration lock on
#     N=30 / tiers=(16384,) / timeout=900s; the CLI defaults still
#     match v2.1, so the locked overrides must be passed explicitly.
#     --bootstrap-n 10000 --permutation-n 10000 aligns the live
#     scorer with score_ext_checkpoints.py):
uv run python -m experiments.v2.p6a.run_live_hf \
    --scope power_ext --live --continue-on-partial \
    --ruler-n 30 --hallu-n 30 --tiers 16384 \
    --scheduler-timeout-s 900 \
    --bootstrap-n 10000 --permutation-n 10000

# (3) offline re-score the checkpoints (no CLI traffic):
uv run python -m tools.dev.score_ext_checkpoints

# (4) consolidator splices the *_score.json overlays in:
uv run python -m tools.dev.live_hf_consolidate

# (5) SWE outcomes side-car (v2.1 + v2.1.1):
uv run python -m tools.dev.sweb_outcomes
\end{lstlisting}

Outputs: \passthrough{\lstinline!.cache/live\_hf\_checkpoints/H\{1\_ruler\_16384,\_TQA\_live\_v2,\_SWEB\}\_live\_ext.jsonl!}
(checkpoints), \passthrough{\lstinline!.cache/live\_hf\_checkpoints/*\_ext\_score.json!}
(offline score side-cars), \passthrough{\lstinline!experiments/results/p6a/\_summary\_live.json!}
(7 bundles: the original 4 v2.1 bundles plus the 3 v2.1.1 \passthrough{\lstinline!\_ext!}
bundles, each carrying pooled + per-model outcomes and caveats),
\passthrough{\lstinline!experiments/results/p6a/swe\_bench\_outcomes.json!}
(v2.1 haiku-only paired slice), \passthrough{\lstinline!experiments/results/p6a/swe\_bench\_outcomes\_ext.json!}
(v2.1.1 haiku-only paired slice, essentially uninformative at $n_\text{pair}=1$).
