\section{Theoretical Foundations: Classical Indian Epistemology Mapped to LLM Context Engineering}\label{theoretical-foundations-classical-indian-epistemology-mapped-to-llm-context-engineering}

This section constructs a precise translation table from seven classical Indian epistemic constructs --- drawn from four philosophical schools, \textbf{Nyāya--Vaiśeṣika}, \textbf{Advaita Vedānta}, \textbf{Pūrva Mīmāṃsā}, and \textbf{Sāṃkhya} --- to seven runtime operations on an LLM agent's context. Each row is a falsifiable design commitment, and each is operationalised in Section 5 and tested in Sections 8--10. To preempt the most common misreading: we use ``classical Indian epistemology'' rather than ``Vedic epistemology'' because the constructs we draw on (the Nyāya \emph{avacchedaka}, the Advaita \emph{bādha}, the Sāṃkhya \emph{manas}/\emph{buddhi} split, the cross-school \emph{khyātivāda} debate) are products of the \emph{darśana} literature, \emph{not} of the Vedic \emph{saṃhitā} corpus per se. Section 4.9 returns to what is and is not being claimed.

\subsection{Pratyakṣa --- direct perception as the harness's grounding pramāṇa}\label{pratyakux1e63a-direct-perception-as-the-harnesss-grounding-pramux101ux1e47a}

In Nyāya \citep{aksapada200nyayasutra, matilal1986perception, phillips2012epistemology} and Advaita \citep{dharmaraja17vedantaparibhasa, datta1932advaita}, \emph{pratyakṣa} (direct perception) is the \emph{first} of the \emph{pramāṇas} --- the means of valid cognition --- and the \emph{foundational} one against which inference (\emph{anumāna}), comparison (\emph{upamāna}), testimony (\emph{śabda}), implication (\emph{arthāpatti}), and non-cognition (\emph{anupalabdhi}) are measured.

\textbf{LLM operationalization.} We treat \emph{pratyakṣa} as the agent's \textbf{direct read of its current context window} --- what it can in-fact attend to right now --- and demote all upstream operations (RAG retrieval, tool outputs, prior-turn summaries) to the status of \emph{secondary pramāṇas} whose claims must be entered into the visible context with their own qualifiers. The harness's name, \emph{Pratyakṣa}, is therefore a methodological commitment: \emph{if it is not in the visible, witness-tracked context, it is not direct evidence for the agent's next action}. Hidden hot caches, opaque rerankers, and ``system memories'' the agent cannot inspect are, in this discipline, \textbf{not allowed to ground a claim} without first entering the \emph{pratyakṣa} surface.

\subsection{Avacchedaka --- qualifier-conditioned cognition as typed limitor}\label{avacchedaka-qualifier-conditioned-cognition-as-typed-limitor}

In Navya-Nyāya \citep{matilal1968navyanyaya, ingalls1951materials, gangesa14tattvacintamani, phillips2012epistemology}, every cognition is structurally a triple:

\[
C = \langle a,\; Q,\; R \rangle \quad\text{read as ``}a\text{ as qualified by }Q\text{ under relation/condition }R\text{''.}
\]

Here \(a\) is the \emph{qualificand} (viśeṣya), \(Q\) the \emph{qualifier} (prakāra), and \(R\) the \emph{limitor} (avacchedaka) under which the qualification holds. Crucially, \emph{R} is part of the cognition; ``the pot is blue \emph{under daylight}'' is a different cognition from ``the pot is blue \emph{under candlelight}'', and the two can co-exist without contradiction.

\textbf{LLM operationalization.} Every item entered into the harness's \passthrough{\lstinline!ContextStore!} carries the four-tuple

\[
\texttt{Item} = \big\langle \texttt{qualificand},\; \texttt{qualifier},\; \texttt{condition},\; \texttt{precision}\in[0,1] \big\rangle.
\]

\passthrough{\lstinline!condition!} is a free-text string like \passthrough{\lstinline!"as of 2024-09-Django-5.0"!} or \passthrough{\lstinline!"on POSIX, Python ≥ 3.11"!}. The MCP tool \passthrough{\lstinline!context\_insert(qualificand, qualifier, condition, precision, source)!} is the only way evidence enters the store. Two items that share \passthrough{\lstinline!qualificand!} but differ in \passthrough{\lstinline!condition!} \emph{do not conflict}; two items that share \passthrough{\lstinline!qualificand!} \emph{and} \passthrough{\lstinline!condition!} but disagree on \passthrough{\lstinline!qualifier!} \emph{do}. This is the operational definition of a contradiction the harness can act on (Section 4.3). Figure \ref{fig:avacchedaka} makes this concrete with two version-typed Django claims that share a qualificand and a qualifier-slot but coexist truthfully because their conditions differ.

\input{figures_tikz/fig2_avacchedaka.tex}

\subsection{Bādha --- sublation as supersede-with-provenance}\label{bux101dha-sublation-as-supersede-with-provenance}

In Advaita Vedānta \citep{deutsch1969advaita, shaw1990bada, dharmaraja17vedantaparibhasa, sankaraupadesha}, a higher-precision cognition \emph{sublates} (bādhita) a lower-precision one. The classical example is mistaking a rope for a snake: the \emph{snake-cognition} is not erased on closer inspection --- the agent must remember it, because \emph{that error is itself an object of subsequent reflection} --- but its qualifier is now stamped \emph{bādhita} and it no longer grounds action.

\textbf{LLM operationalization.} The MCP tool \passthrough{\lstinline!sublate\_with\_evidence(target\_id, by\_id, reason)!} rewrites the target item's \passthrough{\lstinline!status!} to \passthrough{\lstinline!bādhita!}, annotates it with a pointer to the superseding item, and \emph{retains} the original under that status. Subsequent retrievals filter by \passthrough{\lstinline!status == "live"!} for normal use but can be opted-in for audit. Sublation is triggered by two rules: (i) an \textbf{explicit pointer}, when the newly-arriving item names a prior item via \passthrough{\lstinline!superseded\_by\_id!}; and (ii) the \textbf{dominance rule}, when the newly-arriving item shares \passthrough{\lstinline!qualificand!} and \passthrough{\lstinline!condition!} with an existing item, has \emph{strictly higher} precision, and the older item is \passthrough{\lstinline!stale=True!} while the newer is not. The dominance rule is the operational form of the Advaita commitment that \emph{higher pramāṇa supersedes lower pramāṇa} under shared limitor; the explicit pointer is the form used when an LLM-side classifier or external evidence has detected a stale-fresh pair. Figure \ref{fig:sublation} traces the three-step \passthrough{\lstinline!detect\_conflict → gather sources → sublate\_with\_evidence!} procedure end-to-end, including the immutable line written by the Sākṣī Keeper.

\input{figures_tikz/fig3_sublation.tex}

\subsection{Buddhi and Manas --- the two-stage attend-then-judge gate}\label{buddhi-and-manas-the-two-stage-attend-then-judge-gate}

The Antaḥkaraṇa quartet of \emph{manas, buddhi, citta, ahaṃkāra} \citep{deutsch1969advaita, ramprasad2013advaita, datta1932advaita} distinguishes the \emph{attentional sense-organ} (\emph{manas}) from the \emph{determinative judging faculty} (\emph{buddhi}). Manas selects and presents; Buddhi decides. The two are sequential and cannot be collapsed without losing the gate.

\textbf{LLM operationalization.} The harness defines two sub-agent prompts (Section 6.3). \textbf{\passthrough{\lstinline!ManasAgent!}} receives the user query and the \passthrough{\lstinline!ContextStore!} snapshot and returns a structured JSON object matching the shipped contract in \passthrough{\lstinline!agents/manas.md!}: \passthrough{\lstinline!\{ "draft": "...", "grounding": ["<element\_id>", ...], "uncertain\_claims": ["..."], "needs\_buddhi": true | false \}!}; Manas is operationalised as a \textbf{sub-agent} (Section 6.3), not a skill, and is \emph{forbidden} from emitting a final user-visible answer (its draft text is for the orchestrator only). \textbf{\passthrough{\lstinline!BuddhiAgent!}} receives the user query, the Manas output, and re-reads the attended items from the \passthrough{\lstinline!ContextStore!}, then returns the final answer plus its own \passthrough{\lstinline!khyāti\_class!} (Section 4.6); Buddhi is the only agent that may emit a user-visible answer.

This two-stage gate is identical in shape to dual-process accounts of human cognition \citep{evans2003duality, kahneman2011thinking, sloman1996two} and to classic AI cognitive architectures \citep{laird1987soar, anderson1996actr, newell1990unified}, but the names --- and the surrounding philosophical apparatus --- are imported from \textbf{Sāṃkhya} (the original \emph{manas}/\emph{buddhi} enumeration in the \emph{tattva} scheme), as cross-mapped in \textbf{Advaita Vedānta} \citep{deutsch1969advaita, datta1932advaita, ramprasad2013advaita}. Critically, \emph{Manas can be wrong without Buddhi being wrong}: if Manas mis-selects, Buddhi's job is to detect and call for re-selection rather than to answer. Figure \ref{fig:manas_buddhi} shows the resulting two-stage loop and how the Sākṣī Keeper observes both stages without entering the response path.

\input{figures_tikz/fig4_manas_buddhi.tex}

\subsection{Sākṣī --- the witness as model-invariant audit frame}\label{sux101kux1e63ux12b-the-witness-as-model-invariant-audit-frame}

In Advaita Vedānta \citep{indich1980consciousness, fasching2009witness, ganeri2017concealed, sankaraupadesha}, \emph{sākṣī} --- the witness consciousness --- is the \emph{non-revisable, non-acting} observer that \emph{records} what is cognised without itself being the agent of cognition or action. It supplies the stable reference frame against which changes of cognition are intelligible at all.

\textbf{LLM operationalization.} The harness instantiates a \passthrough{\lstinline!SakshiKeeperAgent!} (Section 6.3) and a \passthrough{\lstinline!SakshiPrefix!} skill (Section 6.2). On every turn the keeper appends to a persistent JSON-lines audit log at \passthrough{\lstinline!\~/.cache/pratyaksha/audit.jsonl!} (XDG-style cache path; survives plugin reinstall) an immutable record containing the user query; the Manas selection (which items, under which conditions); the Buddhi judgment (final answer, khyāti class, calibrated posterior); all sublations fired this turn (\passthrough{\lstinline!target\_id!}, \passthrough{\lstinline!by\_id!}, \passthrough{\lstinline!reason!}); and the token budget consumed, broken down by category (system, retrieval, tool, completion).

Crucially, the witness is \emph{write-once-per-turn} and \emph{read-only} to all agents in the next turn. It is the harness's source of truth for cross-turn provenance, and it is \emph{invariant under model swaps}: the witness log written under Claude 3.5 Sonnet is fully readable, fully diffable, and fully auditable when the agent is later switched to Claude 4.5 Sonnet, GPT-4o, or Qwen-3-72B. The model is replaceable; the witness is not.

\subsection{Khyātivāda --- six-class typology of error}\label{khyux101tivux101da-six-class-typology-of-error}

Six classical schools of Indian philosophy proposed six theories of erroneous cognition \citep{ram2007error, matilal1986perception, datta1932advaita, bhatt1989prabhakara, bilimoria2018epistemology, mohanty1992reason}. Each names a specific \emph{kind} of error:

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Class
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
School
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Kernel definition
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Modern LLM analogue
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Anyathākhyāti} & Nyāya & ``Apprehension of one thing as another \emph{of the same general type}.'' & Wrong API name (right module, wrong function); citing the right paper but wrong year. \\
\textbf{Ātmakhyāti} & Yogācāra & ``Projection of an internal cognition onto the external object.'' & Inventing helper functions / config keys the framework never exposed; hallucinated TODOs the user did not write. \\
\textbf{Anirvacanīyakhyāti} & Advaita & ``Apprehension of an object that is \emph{neither real nor unreal}; sui generis indeterminate.'' & Hedged hallucinations: ``you might consider X-or-similar'' where X is plausible but ungrounded. \\
\textbf{Asatkhyāti} & Mādhyamika & ``Apprehension of a non-existent object.'' & Citing a CVE, a person, a paper, or a commit hash that simply does not exist. \\
\textbf{Viparītakhyāti} & Mīmāṃsā & ``Apprehension of the contrary.'' & Inverting a boolean flag's semantics; recommending the \emph{opposite} of the documented best practice. \\
\textbf{Akhyāti} & Prabhākara Mīmāṃsā & ``Failure of correct apprehension by mere conflation of two adjacent cognitions.'' & Mixing two adjacent functions / two near-duplicate config schemas / two adjacent versions. \\
\end{longtable}
}

This is not a post-hoc relabelling of an existing modern taxonomy. It is the \emph{original} taxonomy from the Indian philosophical tradition, jointly developed by debating schools over roughly a thousand years. We adopt it directly because (a) it carves the space cleanly along \emph{epistemic} (not merely surface-textual) lines, (b) the six classes are mutually exclusive in the tradition's own usage, and (c) it gives us a \emph{typed signal} the harness can act on (different remediation strategies for different classes).

\textbf{LLM operationalization.} The shipped plugin's \passthrough{\lstinline!classify\_khyativada!} tool applies a \textbf{heuristic classifier with the rule-based guardrails} described in Section 5.5. Separately, the experiments harness includes a few-shot Anthropic JSON classifier (\passthrough{\lstinline!src/evaluation/khyativada\_fewshot.py!}) that is \textbf{not} wired into the plugin tool; both paths are evaluated independently in §8 (H6). Inter-annotator agreement is reported on the 3,000-example corpus using prompts in Appendix D and results in §8.6 at Cohen's κ = 0.736 (``substantial''). Figure \ref{fig:khyati} summarises the six-class taxonomy and gives one canonical agentic-AI failure example per class.

\input{figures_tikz/fig5_khyati.tex}

\subsection{Saṃskāras and vāsanās --- adaptive forgetting}\label{saux1e43skux101ras-and-vux101sanux101s-adaptive-forgetting}

In Advaita and Yoga, \emph{saṃskāras} are residual mental impressions and \emph{vāsanās} are dispositional tendencies that survive the immediate occasion of a cognition \citep{deutsch1969advaita, sankaraupadesha, halbfass1991traditioncomparison}. They explain why we are drawn to the same patterns of thought and why some impressions decay quickly while others persist indefinitely. The Advaitin practice prescribes \emph{removal of obstructive vāsanās} --- adaptive forgetting of patterns that no longer ground correct cognition.

\textbf{LLM operationalization.} The \passthrough{\lstinline!AdaptiveForgetting!} module (Section 5.7) applies an exponential-decay schedule to non-witnessed items in the \passthrough{\lstinline!ContextStore!}, governed by three explicit rules. \textbf{Witnessed items never decay}: if \passthrough{\lstinline!sākṣī.witnessed == True!} for an item-id, its precision floor is fixed. \textbf{Sublated items decay faster}: items with \passthrough{\lstinline!status == bādhita!} halve their precision every time they survive a budget-pressure compaction event. \textbf{High-recency, low-precision items decay slowest among the rest} --- the item is too new to know whether it will become a vāsanā or a saṃskāra. This avoids both the catastrophic-forgetting failure mode of naive context pruning \citep{french1999catastrophic, kirkpatrick2017overcoming, parisi2019continual} and the ``everything is forever'' failure mode of unconstrained external memory \citep{packer2023memgpt}.

\subsection{The translation table, summarised}\label{the-translation-table-summarised}

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Construct (school)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
LLM operationalisation
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
MCP tool / module
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Tested by
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\emph{Pratyakṣa} (Nyāya, Advaita) & Visible-context-only grounding discipline & \passthrough{\lstinline!context\_retrieve!}, \passthrough{\lstinline!context\_window!}, related read paths & All hypotheses (architectural commitment) \\
\emph{Avacchedaka} (Navya-Nyāya) & Typed limitor \passthrough{\lstinline!(qualificand, qualifier, condition, precision)!} & \passthrough{\lstinline!context\_insert!}, \passthrough{\lstinline!context\_retrieve!} & H2, H5 \\
\emph{Bādha} (Advaita Vedānta + Pūrva Mīmāṃsā) & Supersede-with-provenance under shared limitor & \passthrough{\lstinline!sublate\_with\_evidence!} & H4, H5, P6-B, P6-C \\
\emph{Manas} (Sāṃkhya, Advaita) & Pre-judgment attention-selection \textbf{sub-agent} & \passthrough{\lstinline!ManasAgent!} (Section 6.3) & H3 \\
\emph{Buddhi} (Sāṃkhya, Advaita) & Determinative judgment + \emph{khyāti} tagging sub-agent & \passthrough{\lstinline!BuddhiAgent!} & H3, H6 \\
\emph{Sākṣī} (Advaita Vedānta) & Write-once cross-turn witness log & \passthrough{\lstinline!SakshiKeeperAgent!}, \passthrough{\lstinline!SakshiPrefix!} & All (audit invariant) \\
\emph{Khyātivāda} (cross-school: Nyāya, Yogācāra, Advaita, Madhyamaka, Mīmāṃsā, Prabhākara, Viśiṣṭādvaita) & 6-class hallucination classifier & \passthrough{\lstinline!classify\_khyativada!} & H6 \\
\emph{Saṃskāra} / \emph{vāsanā} (Advaita / Yoga) & Adaptive forgetting with witness-protected items & \passthrough{\lstinline!AdaptiveForgetting!} & H7 \\
\end{longtable}
}

\subsection{\texorpdfstring{What is \emph{not} being claimed}{What is not being claimed}}\label{what-is-not-being-claimed}

We are \emph{not} claiming that the historical authors had LLMs in mind, nor that our operationalizations are textually faithful in the way a Sanskrit philologist would require. We are claiming the weaker, falsifiable thing: that the \emph{type signatures} the tradition developed for cognition turn out to \emph{fit} the type signatures we need for an LLM context-engineering harness, and that fitting them produces measurable downstream gains. The next six sections defend that claim empirically.
