\section{6 · The Cursor / Claude-Code Plugin}\label{the-cursor-claude-code-plugin}

The harness ships as a single plugin, \textbf{\passthrough{\lstinline!pratyaksha-context-eng-harness!}} (v1.0.0, MIT-licensed), installable into Cursor, Claude Code (CLI and the VS Code extension), and Claude Desktop via the Model Context Protocol \citep{anthropic2025mcp}. This section specifies the shipped artefact in enough detail that an external reviewer can verify its claims directly against the repository at \passthrough{\lstinline!github.com/SharathSPhD/pratyaksha-context-eng-harness!}.

\subsection{6.1 Manifest and discovery}\label{manifest-and-discovery}

Two manifests are shipped:

\begin{itemize}
\tightlist
\item
  \passthrough{\lstinline!.claude-plugin/plugin.json!} --- Claude-Code-format component manifest declaring the 3 agents, 4 commands, 3 skills, and the \passthrough{\lstinline!hooks/hooks.json!} registration.
\item
  \passthrough{\lstinline!marketplace.json!} --- top-level marketplace manifest used for \passthrough{\lstinline!/plugin install!} discovery, listing the canonical id, version, repository URL, license, keywords, and feature catalogue.
\end{itemize}

Both manifests are version-pinned to \passthrough{\lstinline!1.0.0!}, point to the same upstream URL, and resolve identically in Cursor (via the plugin marketplace \citep{cursor2025plugins}) and Claude Code (via the new plugin marketplace surface \citep{anthropic2025claudecode}). Setup is a single command: \passthrough{\lstinline!curl -LsSf https://astral.sh/uv/install.sh | sh!}; the MCP server's Python dependencies are installed lazily on first tool call by \passthrough{\lstinline!uv!} \citep{uv2025}, so a fresh user can install the plugin without ever invoking \passthrough{\lstinline!pip!}, creating a virtualenv, or running \passthrough{\lstinline!claude mcp add!}.

\subsection{6.2 The 15 MCP tools}\label{the-15-mcp-tools}

The MCP server (\passthrough{\lstinline!mcp/server.py!}, FastMCP-style) exposes exactly fifteen tools. Their names, signatures, and operational roles are:

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\#
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Tool
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Signature
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Operationalizes
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1 & \passthrough{\lstinline!context\_insert!} & \passthrough{\lstinline!InsertInput → ContextElement!} & Avacchedaka insertion (Section 4.2) \\
2 & \passthrough{\lstinline!context\_retrieve!} & \passthrough{\lstinline!RetrieveInput → list[ContextElement]!} & Default avacchedaka query \\
3 & \passthrough{\lstinline!context\_get!} & \passthrough{\lstinline!GetByIdInput → ContextElement!} & Direct id-based fetch (audit) \\
4 & \passthrough{\lstinline!context\_sublate!} & \passthrough{\lstinline!SublateInput → ContextElement!} & Bādha (basic, by id) \\
5 & \passthrough{\lstinline!sublate\_with\_evidence!} & \passthrough{\lstinline!SublateWithEvidenceInput → ...!} & Bādha with explicit pointer + reason \\
6 & \passthrough{\lstinline!detect\_conflict!} & \passthrough{\lstinline!RetrieveInput → \{conflicted: bool, items: ...\}!} & Bayesian-margin conflict check \\
7 & \passthrough{\lstinline!list\_qualificands!} & \passthrough{\lstinline!() → list[str]!} & Diagnostic surface \\
8 & \passthrough{\lstinline!compact!} & \passthrough{\lstinline!CompactInput → CompactionReport!} & Adaptive forgetting (Section 5.7) \\
9 & \passthrough{\lstinline!boundary\_compact!} & \passthrough{\lstinline!BoundaryCompactInput → ...!} & Event-boundary compaction (Section 5.6) \\
10 & \passthrough{\lstinline!context\_window!} & \passthrough{\lstinline!ContextWindowInput → WindowSnapshot!} & Visible-context summary for Manas \\
11 & \passthrough{\lstinline!set\_sakshi!} & \passthrough{\lstinline!SetSakshiInput → ()!} & Pin a session-stable witness invariant \\
12 & \passthrough{\lstinline!get\_sakshi!} & \passthrough{\lstinline!() → list[SakshiInvariant]!} & Retrieve the witness invariants \\
13 & \passthrough{\lstinline!classify\_khyativada!} & \passthrough{\lstinline!ClassifyInput → KhyatiResult!} & 7-class hallucination classifier \\
14 & \passthrough{\lstinline!budget\_status!} & \passthrough{\lstinline!BudgetStatusInput → BudgetStatus!} & TokenBudgetWatchdog (Section 5.8) \\
15 & \passthrough{\lstinline!budget\_record!} & \passthrough{\lstinline!BudgetRecordInput → BudgetStatus!} & Record consumed tokens per turn \\
\end{longtable}
}

Every tool is typed via Pydantic v2 input/output models, returns JSON-serializable structures, and is exercised by at least one unit test in \passthrough{\lstinline!tests/test\_v2/!}. A complete tool reference with full schemas is in \textbf{Appendix B}.

\subsection{6.3 The 3 sub-agents}\label{the-3-sub-agents}

\begin{itemize}
\tightlist
\item
  \textbf{\passthrough{\lstinline!agents/manas.md!}} --- defines the \emph{attentional sense-organ} \textbf{sub-agent}. System prompt instructs the model to call \passthrough{\lstinline!get\_sakshi!}, \passthrough{\lstinline!context\_retrieve!}, \passthrough{\lstinline!context\_window!}, \passthrough{\lstinline!classify\_khyativada!}, and \passthrough{\lstinline!budget\_record!}, and to return structured JSON of the form \passthrough{\lstinline!\{ "draft": "...", "grounding": ["<element\_id>", ...], "uncertain\_claims": ["..."], "needs\_buddhi": true | false \}!} (verbatim contract in the repository file). Manas must \emph{never} emit a user-visible final answer.
\item
  \textbf{\passthrough{\lstinline!agents/buddhi.md!}} --- the \emph{determinative judging} sub-agent. May call \passthrough{\lstinline!sublate\_with\_evidence!}, \passthrough{\lstinline!detect\_conflict!}, and \passthrough{\lstinline!classify\_khyativada!}. Emits the user-visible answer and its self-classified \passthrough{\lstinline!khyati\_class!} and \passthrough{\lstinline!confidence!}.
\item
  \textbf{\passthrough{\lstinline!agents/sakshi-keeper.md!}} --- the \emph{witness keeper}. Read-only on the live store; append-only on the JSON-lines audit log at \passthrough{\lstinline!\~/.cache/pratyaksha/audit.jsonl!}. Invoked from the lifecycle hooks (Section 6.5).
\end{itemize}

The choice to ship Manas and Buddhi as \emph{agents} (system-prompt-only sub-agents driven by the host LLM) rather than as fine-tunes is deliberate: it preserves the harness's host-platform-agnosticism (Section 5.4) and lets users swap the underlying model without re-training anything.

\subsection{6.4 The 3 skills}\label{the-3-skills}

Following the Cursor/Claude-Code skill format (an \passthrough{\lstinline!SKILL.md!} per directory):

\begin{itemize}
\tightlist
\item
  \textbf{\passthrough{\lstinline!skills/context-discipline/SKILL.md!}} --- surfaces the harness's discipline as a triggered skill: \emph{``every retrieved item must be entered with avacchedaka qualifiers; no free-text dumps''}. Activates when the host agent is about to call a search/RAG tool.
\item
  \textbf{\passthrough{\lstinline!skills/sublate-on-conflict/SKILL.md!}} --- triggers Buddhi's call to \passthrough{\lstinline!sublate\_with\_evidence!} when \passthrough{\lstinline!detect\_conflict!} returns true.
\item
  \textbf{\passthrough{\lstinline!skills/witness-prefix/SKILL.md!}} --- wraps every Buddhi prompt with a stable \passthrough{\lstinline!<sakshi\_invariants>!} system block via \passthrough{\lstinline!get\_sakshi!}, so the witness frame is \emph{literally a system message}, not inlined into the user turn.
\end{itemize}

\subsection{6.5 The 4 slash commands}\label{the-4-slash-commands}

\begin{itemize}
\tightlist
\item
  \textbf{\passthrough{\lstinline!/context-status!}} --- pretty-prints the visible context window, by category, with budget gauge.
\item
  \textbf{\passthrough{\lstinline!/sublate <target\_id> <by\_id> <reason>!}} --- manual sublation override (audit / debug).
\item
  \textbf{\passthrough{\lstinline!/budget!}} --- shorthand for \passthrough{\lstinline!budget\_status!} rendered as a one-line gauge.
\item
  \textbf{\passthrough{\lstinline!/compact-now [strategy]!}} --- manually trigger compaction; \passthrough{\lstinline!strategy ∈ \{adaptive, lru, none\}!} (default \passthrough{\lstinline!adaptive!}).
\end{itemize}

\subsection{6.6 The 3 lifecycle hooks}\label{the-3-lifecycle-hooks}

\passthrough{\lstinline!hooks/hooks.json!} registers three hooks against the standard Claude Code lifecycle events:

\begin{lstlisting}
{
  "hooks": {
    "SessionStart": [{ "matcher": "*", "hooks": [
      { "type": "command", "command": "${CLAUDE_PLUGIN_ROOT}/hooks/session-start.sh", "timeout": 10 }
    ]}],
    "PreToolUse": [{ "matcher": "mcp__pratyaksha_mcp__.*", "hooks": [
      { "type": "command", "command": "${CLAUDE_PLUGIN_ROOT}/hooks/pretooluse-budget.sh", "timeout": 5 }
    ]}],
    "Stop": [{ "matcher": "*", "hooks": [
      { "type": "command", "command": "${CLAUDE_PLUGIN_ROOT}/hooks/stop-compact.sh", "timeout": 5 }
    ]}]
  }
}
\end{lstlisting}

Their roles are:

\begin{itemize}
\tightlist
\item
  \textbf{\passthrough{\lstinline!session-start.sh!}} seeds the session-stable Sākṣī invariants (working directory, git SHA, model id, plugin version) via \passthrough{\lstinline!set\_sakshi!}.
\item
  \textbf{\passthrough{\lstinline!pretooluse-budget.sh!}} consults \passthrough{\lstinline!budget\_status!} before any harness tool runs and \textbf{logs} over-threshold use by default (advisory hook). If \passthrough{\lstinline!PRATYAKSHA\_BUDGET\_STRICT=1!}, the same script can be configured to \textbf{deny} tool execution when the gauge exceeds a hard threshold (default 95\% of \passthrough{\lstinline!context\_budget!}).
\item
  \textbf{\passthrough{\lstinline!stop-compact.sh!}} invokes \passthrough{\lstinline!compact(strategy="adaptive")!} at the end of every turn so memory pressure does not accumulate across turns.
\end{itemize}

\subsection{6.7 Worked example: a Redis-caching turn, end-to-end}\label{worked-example-a-redis-caching-turn-end-to-end}

To make the runtime contract concrete we trace a single user turn end-to-end through the deployed plugin. The user prompt is \emph{``how do I cache a user session in Redis?''}; the agent's tool returns a mix of pre-Redis-7 blog snippets and the official Redis 7 documentation. Figure\textasciitilde{}\ref{fig:swimlane} visualises the swimlane across \textbf{User → Manas → Buddhi → Sublation → Sākṣī}, with the per-stage immutable JSON line written into \passthrough{\lstinline!\~/.cache/pratyaksha/audit.jsonl!}.

\input{figures_tikz/fig6_swimlane.tex}

The five host-visible artefacts are: (i) one \passthrough{\lstinline!mcp\_\_pratyaksha\_mcp\_\_manas\_step!} JSON-RPC call (Manas attends, \(K=3\) items under a 1.2 K-token budget); (ii) one \passthrough{\lstinline!detect\_conflict!} round-trip flagging a TYPE\_CLASH on the \passthrough{\lstinline!(Redis-session, expiry-policy)!} qualificand-qualifier slot; (iii) one \passthrough{\lstinline!sublate\_with\_evidence!} call that retires the blog post in favour of the Redis 7 docs (limitor precedence: official docs \passthrough{\lstinline!prec=8!} \textgreater{} blog post \passthrough{\lstinline!prec=2!}); (iv) one \passthrough{\lstinline!mcp\_\_pratyaksha\_mcp\_\_buddhi\_step!} call returning the answer with \passthrough{\lstinline!khyati\_class = "yathārtha"!} (veridical) and \passthrough{\lstinline!confidence = 0.91!}; and (v) one append to the Sākṣī log per stage. The complete \passthrough{\lstinline!/context-status!} snapshot, the raw JSON-RPC payloads, and the four matching audit-log lines are reproduced verbatim in \textbf{Appendix C.8} so that the turn can be replayed byte-for-byte from the shipped plugin against the cached evidence trail.

This single turn exercises every load-bearing primitive of Sections 4--5 in a non-coding agentic context, and is the runtime template the L1 (§8) and L3 (§10) experiments instantiate at scale.

\subsection{\texorpdfstring{6.8 What the plugin does \emph{not} contain}{6.8 What the plugin does not contain}}\label{what-the-plugin-does-not-contain}

We assert and audit two negative claims:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Zero references to the dev-time orchestration tools} \passthrough{\lstinline!attractor-flow!}, \passthrough{\lstinline!ralph-loop!}, and \passthrough{\lstinline!triz-engine!}. Those tools were used to build the harness (Section 3) but are not runtime dependencies. A grep over the entire shipped plugin tree returns zero matches; the audit is preserved as \passthrough{\lstinline!docs/no\_dev\_deps\_audit.md!} in the repository.
\item
  \textbf{Zero hard dependencies on heavyweight stacks} like \passthrough{\lstinline!vllm!}, \passthrough{\lstinline!mlflow!}, \passthrough{\lstinline!chromadb!}, or \passthrough{\lstinline!huggingface-hub!}. The harness's runtime requirements are \passthrough{\lstinline!mcp!}, \passthrough{\lstinline!pydantic!}, \passthrough{\lstinline!tiktoken!}, \passthrough{\lstinline!numpy!}, and \passthrough{\lstinline!anthropic!}; the optional Qwen3-1.7B surprise backend is loaded only when the user opts in.
\end{enumerate}

This matters because the harness is a \emph{discipline}, not an infrastructure. Anything heavier than that would defeat the purpose of a plugin you can install in 30 seconds.

\subsection{6.9 Smoke test and CI}\label{smoke-test-and-ci}

\passthrough{\lstinline!mcp/smoke\_test.py!} exercises every tool against the local MCP server: 15 round-trips, each verifying the input schema, the output schema, and one non-trivial behavioural assertion (e.g.~\passthrough{\lstinline!sublate\_with\_evidence!} actually flips the target item's status). The smoke test runs in roughly 4 seconds wall-clock and is wired into the repo's CI as \passthrough{\lstinline!pytest -m smoke!}. We additionally exercise every command and hook via \passthrough{\lstinline!claude --debug!} in the developer's local Claude Code installation; the transcript of one such smoke run is preserved in \passthrough{\lstinline!docs/plugin\_smoke\_transcript.md!}.

\subsection{6.10 Hot-swappability across hosts}\label{hot-swappability-across-hosts}

The same \passthrough{\lstinline!marketplace.json!} resolves in:

\begin{itemize}
\tightlist
\item
  \textbf{Cursor} --- via the IDE's plugin browser, registered by URL.
\item
  \textbf{Claude Code (CLI)} --- via \passthrough{\lstinline!/plugin install <url>!}.
\item
  \textbf{Claude Code (VS Code extension)} --- via the same \passthrough{\lstinline!/plugin install!} UI.
\item
  \textbf{Claude Desktop} --- via the Desktop app's MCP-based plugin surface.
\end{itemize}

We have manually exercised the install path on all four hosts during development. The harness's runtime invariants (Sections 4 and 5) hold identically across all four because the only inter-process surface is MCP, and MCP is host-agnostic by construction \citep{anthropic2025mcp}.
