\section{Discussion, Limitations, and Threats to Validity}\label{discussion-limitations-and-threats-to-validity}

We have presented a context-engineering system whose mechanisms are imported from classical Indian epistemology --- specifically Nyāya--Vaiśeṣika, Advaita Vedānta, Pūrva Mīmāṃsā, and Sāṃkhya --- and whose validation produces a Stouffer-combined \(p = 7.94 \times 10^{-20}\). The reader is rightly suspicious of any number that small. This section names --- and where possible bounds --- the threats that should temper it, and discusses what the system genuinely contributes once those threats are accounted for.

\subsection{Threats to validity, named}\label{threats-to-validity-named}

\textbf{Synthetic-fallback adapter risk (L1).} Most of the seven L1 hypotheses (H1--H7) use the \emph{synthetic-fallback} path of their adapter (Wikipedia + arXiv distractors are loaded only when an HF token is configured; the path actually run in CI uses deterministic synthetic generators). This was a deliberate trade-off: it lets the entire validation re-run on a laptop in under three minutes for full reproducibility, but it shifts the burden of ``is this benchmark \emph{like} the published one?'' onto our generator. We mitigated by (i) tuning generator parameters so per-context-length difficulty curves qualitatively matched the published RULER and HELMET curves \citep{hsieh2024ruler, yen2024helmet}, and (ii) running the \emph{real} HF-loaded path during P3 development to verify that treatment/baseline gaps were within ±10\% of the synthetic-fallback path on a 3,000-example sample. We do not claim our numbers are \emph{the same as} published RULER / HELMET numbers --- only that the \emph{paired delta in the system's favour} is a faithful estimate.

\textbf{Live Hugging Face rerun (v2.1, \texttt{core4}; and v2.1.1 power-extension addendum).} To put a partial floor under the synthetic-fallback threat we ran a pre-registered, \passthrough{\lstinline!strict\_hf!}-gated \textbf{live} pass on four headline bundles against pinned commit SHAs of \passthrough{\lstinline!simonjegou/ruler!}, \passthrough{\lstinline!truthfulqa/truthful\_qa!}, and \passthrough{\lstinline!princeton-nlp/SWE-bench\_Verified!} (resolved at consolidation time via the Hub API and stamped into \passthrough{\lstinline!provenance.hf\_revision!} on every record) (Appendix~\ref{live-hf-rerun-v2.1}, Table~\ref{tab:t8_live_hf}). At the $N=60$/$N=30$ paired-observation budget that a single Anthropic rolling window affords, the rerun \emph{reproduces} the RULER 8\,K effect ($\Delta=+0.233$, $p=0.0005$, $d_z=0.547$), shows the RULER 16\,K effect only \emph{directionally} (underpowered), returns a \emph{null} on TruthfulQA, and reproduces the SWE-bench direction at a smaller effect on the haiku-only slice that completed ($\Delta=+0.109$, $p=0.032$, $d_z=0.488$). A follow-on \textbf{v2.1.1 power-extension amendment} (Appendix~\ref{live-hf-v2.1.1}) doubled $N$ from 15 to 30 on RULER 16\,K and TruthfulQA and raised the SWE-bench CLI timeout to eliminate the \passthrough{\lstinline!SessionStart!}-hook abort path; Anthropic's account-level rolling quota exhausted partway through, and we halted the run. The RULER 16\,K read at $n_\text{pair}=103$ \emph{tightens the interval} --- 95\% bootstrap CI now \textbf{[+0.010,\,+0.087] excluding zero} --- while falling just short of both preregistered gates ($\Delta=+0.0485$, $p=0.064$, $d_z=0.225$). TruthfulQA received no new rows and its null persists at $N=60$; the SWE-bench CLI fix shipped but was not exercised under load. We therefore narrow the synthetic-fallback claim: the paired delta is a faithful estimate \emph{in direction and easy-end magnitude}, but the live rerun cannot, at this $N$, corroborate a specific per-benchmark effect size on TruthfulQA or RULER 16\,K, and the SWE-bench live slice remains partially CLI-blocked (Appendix~\ref{live-hf-sweb-cli-limitation}). The conservative reading across both passes together is that the system clears \emph{both} preregistered gates ($d \ge 0.5$ \emph{and} $p \le 0.05$) on one independently-sourced live surface (RULER 8\,K, $d_z=0.547$) and the $p \le 0.05$ gate on a second (SWE-bench-haiku, $d_z=0.488$, just below the $d$ gate), shows a direction-consistent but still-underpowered effect on a third (RULER 16\,K at $n_\text{pair}=103$), and returns a null on a fourth (TruthfulQA at $N=60$) --- in addition to clearing the full v2.0 omnibus on synthetic-fallback adapters at $N=180$--$N=700$.

\textbf{PatchSimulator vs.~real LLM coder (L3).} The L3 P6-C study uses a deterministic \passthrough{\lstinline!PatchSimulator!} rather than a real LLM-based code generator. This deliberately isolates the system's contribution as a \emph{context discipline} rather than as generation quality, but it means we \emph{underspecify} the gain a real LLM coder would see. As Section 10.4 notes, a strong LLM coder might recover from some wrong-path anchors that the simulator does not. We predict --- but have not measured --- that the system's gain on a real-LLM-coder version of P6-C will compress to perhaps +0.10 to +0.15 absolute target-path-hit-rate while remaining significant; future work will measure this. Other coder benchmarks \citep{liu2023humanevalplus, austin2021programsynth, chen2021codex, zhuo2024bigcode, wan2024llmsurvey} are natural targets. Crucially, the L1 evidence is unaffected by this threat: it is the \emph{general} claim, and it does not rely on the patch simulator at all.

\textbf{Two model families only.} We sweep \passthrough{\lstinline!claude-haiku-4-5!} and \passthrough{\lstinline!claude-sonnet-4-6!} (Section 7.5), both from Anthropic. The system's design is host- and model-agnostic (Sections 5.4, 6.9), but we have not yet \emph{measured} it across families (GPT-4o-class, Qwen-3, Llama-3.x). The expected confound is \emph{not} that the system will stop working --- its mechanisms are LLM-side prompt discipline plus an MCP-side store, neither of which depends on the model --- but that the \emph{baseline} will be stronger on some families and weaker on others, compressing or expanding the delta. Cross-family validation is a Section 12 priority.

\textbf{Two-rater IAA on Khyātivāda.} The Cohen's \(\kappa = 0.736\) reported in §8.6 is between \emph{two automated annotators}: a deterministic heuristic and a simulated LLM-as-judge built from the few-shot Claude classifier whose prompts are documented in \textbf{Appendix D} (IAA numbers are in §8.6, not Appendix E). A \emph{human-vs-human} IAA on a sample of the same 3,000 examples is the obvious next step and would be the gold standard. Our preliminary read --- by the author, on a 200-example random sample --- agrees with the consensus label in 81\% of cases (\(\kappa \approx 0.74\)), which is consistent with the automated \(\kappa\) but not yet at the scale that would let us claim ``human-validated''. This is the cleanest near-term improvement to make.

\textbf{Two structural-100\% studies inflate the omnibus.} H5 and H7 are \emph{structural} hypotheses: the baseline cannot, by construction, reach the answer that the system reaches. We retain them as sanity checks of the system's \emph{bādha} and adaptive-forgetting implementations rather than as comparative evidence. The Stouffer-Z statistic \emph{excludes} them from the per-Cohen's-d mean (we report \(d = 9.62\) \emph{excluding} them, and an inflated \(d\) if included). The combined \(p\) of \(7.94 \times 10^{-20}\) uses all 10 studies, but we additionally report a ``robust'' \(p\) computed only on the 8 non-structural studies: combined \(Z = \mathbf{8.14}\), two-sided \(p = \mathbf{3.95 \times 10^{-16}}\). The headline still survives by 16 orders of magnitude.

\textbf{Author-as-annotator and author-as-architect.} The author of the paper is the author of the system, the case-study evidence, the Khyātivāda corpus, and the heuristic annotator. This is the standard ``self-as-experimenter'' risk in single-author systems work. Mitigations are that (i) the L1 benchmarks are external (RULER, HELMET, NoCha, HaluEval, TruthfulQA, FACTS, SWE-bench Verified), (ii) the LLM-as-judge annotator is independent of the heuristic annotator, and (iii) all artefacts are open-sourced. The reproducibility manifest (Appendix C) lets a third party re-run the entire pipeline.

\textbf{Construct validity of the philosophical mappings.} We claim that \emph{avacchedaka} (Nyāya--Vaiśeṣika) maps to typed limitor, \emph{bādha} (Advaita Vedānta) to supersede-with-provenance, and so on. A scholar of classical Indian philosophy might reasonably argue that our operationalisations under-respect the technical Sanskrit (e.g.~Navya-Nyāya \emph{avacchedaka} is a far richer relational object than our four-tuple). We accept this --- see Section 4.9 --- and the claim we \emph{do} make is the weaker, falsifiable one: \emph{the operationalisation works, in the sense that it produces measurable downstream gains}. We invite philological critique and welcome refinements that \emph{strengthen} the operationalisation while remaining implementable in a runtime store.

\textbf{Bias from the same author also building \passthrough{\lstinline!triz-engine!} and \passthrough{\lstinline!attractor-flow!}.} Section 3 narrates how this work emerged from a TRIZ session run on the same author's \passthrough{\lstinline!triz-engine!} plugin, with \passthrough{\lstinline!attractor-flow!} driving the long-running build pipeline. We claim --- and audit --- that \emph{neither tool is a runtime dependency of the shipped plugin} (Section 6.7). The methodological dependency on TRIZ is real and is part of the contribution (Section 3.3); the runtime dependency is, by audit, zero.

\subsection{What the system genuinely contributes}\label{what-the-system-genuinely-contributes}

Once the threats are accounted for, the substantive claim narrows but does not vanish:

\begin{quote}
\emph{Under a fixed token budget, on benchmark surfaces designed to expose long-context and hallucination failure modes --- both general (RULER, HELMET, NoCha, HaluEval, TruthfulQA, FACTS-Grounding) and one challenging coding instantiation (SWE-bench Verified) --- an LLM agent equipped with the Pratyakṣa system produces measurably and replicably better answers than the same LLM agent without it, with effect sizes (Cohen's \(d \approx 1.0\) on the most ecologically valid coding study, P6-C) and statistical significance (combined \(p < 10^{-15}\) even on the threat-conservative 8-study subset) far above conventional thresholds.}
\end{quote}

The mechanisms responsible are, in our reading, four interlocking moves. \textbf{Avacchedaka-typed insertion (Nyāya--Vaiśeṣika)} ensures the agent stops asserting facts without conditions, so conflicts become detectable. \textbf{Bādha (sublation) under shared limitor (Advaita Vedānta + Pūrva Mīmāṃsā)} ensures the agent stops carrying sublated evidence into the live answer set, so \emph{Lost-in-the-Middle}-style anchoring is structurally suppressed. \textbf{The Manas/Buddhi two-stage gate (Sāṃkhya, cross-mapped in Advaita)} ensures that \emph{ātmakhyāti}-class hallucinations (projection-of-self) are caught at the \emph{attention} step rather than the \emph{answer} step. \textbf{Sākṣī-tracked cross-turn provenance (Advaita Vedānta)} ensures witness-protected facts survive compaction --- the simplest possible defence against catastrophic forgetting \citep{french1999catastrophic, kirkpatrick2017overcoming, parisi2019continual}.

\subsection{Why the philosophical vocabulary matters at all}\label{why-the-philosophical-vocabulary-matters-at-all}

A reasonable reviewer might ask: granted that the mechanisms work, why import Sanskrit terms from four distinct schools instead of inventing fresh English ones? Three answers.

First, \emph{the mechanisms come as a set}. \emph{Avacchedaka}, \emph{bādha}, \emph{manas}/\emph{buddhi}, \emph{sākṣī}, \emph{khyātivāda}, and \emph{saṃskāra}/\emph{vāsanā} were \emph{jointly developed} by mutually critical philosophical schools over centuries. Picking up any one of them in isolation reproduces the partiality that contemporary context-engineering already suffers from (Section 2.3). The set carries \emph{internal consistency constraints} that we found ourselves needing in any case, and the tradition supplies them.

Second, \emph{the tradition's terms have type signatures the modern field lacks}. ``Sublation'' is not ``delete'' and not ``deprecate''; it is \emph{specifically} ``supersede-with-provenance under shared limitor''. No comparably tight English term exists in the modern literature. Importing the Sanskrit term lets us name what we mean exactly. We attribute each Sanskrit term to its school of origin (Section 4) so the reader can locate the technical sense in the primary literature \citep{matilal1986perception, ganeri2001philosophy, phillips2012epistemology}.

Third, \emph{cross-cultural convergence is suggestive, not yet evidence}. Cognitive neuroscience independently arrived at structurally similar constructs --- working-memory schemas \citep{baddeley1974wm, baddeley2000episodic, ghosh2014schema}, predictive-coding precision \citep{rao1999predictive, friston2010fep}, complementary-learning-systems consolidation \citep{mcclelland1995cls, kumaran2016cls}, prefrontal attention control \citep{miller2001integrative, corbetta2002dorsalventral}, event-segmentation \citep{zacks2007event, baldassano2017nested} --- without ever talking to the Indian tradition. Two unrelated lines of inquiry converging on similar type signatures is \textbf{suggestive convergence}, not by itself proof that those signatures carve reality at a joint; it nevertheless motivates treating the school-attributed packaging as more than mere ornament. We import these terms not for exoticism but because \emph{they were ready first} and are \emph{the integrated form}.

\subsection{Generality across agentic-AI domains}\label{generality-across-agentic-ai-domains}

The mechanisms are not coding-specific. \emph{Avacchedaka}-typed insertion is equally useful for a customer-service agent that must distinguish ``policy as of 2023-Q1'' from ``policy as of 2025-Q4''; for a \textbf{research-assistance} agent that surfaces a paper and then a retraction; for a \textbf{document-QA} agent navigating versioned documentation; for a \textbf{multi-tool orchestrator} integrating outputs from heterogeneous tools whose provenance must remain auditable; for a \textbf{code-review agent} triaging contradictory PR comments; and for any \textbf{long-horizon planning} agent whose intermediate state must survive compaction. \emph{Bādha} is equally useful wherever newer authoritative evidence must displace older evidence without losing the audit trail. \emph{Manas}/\emph{Buddhi} is equally useful in any setting in which selecting evidence and judging on the basis of it are different cognitive acts. The Layer-1 evidence (RULER, HELMET, NoCha, HaluEval, TruthfulQA, FACTS-Grounding) directly exercises these \emph{general} settings; the L2 case study spans Django, Requests, and pandas across three issue types; and the L3 SWE-bench Verified A/B is one specific coding instantiation. We expect --- but have not yet measured at scale --- that the system will generalise to additional non-coding agentic surfaces (legal-document analysis, scientific-literature triage, customer-policy retrieval, agentic search). Cross-domain validation across non-coding agents is the most natural follow-up beyond cross-family LLM validation.
