\begin{center}
\small
\href{https://github.com/SharathSPhD/pratyaksha-context-eng-harness}{\texttt{github.com/SharathSPhD/pratyaksha-context-eng-harness}} \quad\textbar\quad
Plugin marketplace: \href{https://github.com/SharathSPhD/pratyaksha-context-eng-harness/blob/main/.claude-plugin/marketplace.json}{\texttt{.claude-plugin/marketplace.json}} \\
Reproducibility manifest: \texttt{experiments/results/p7/\_index.json}, \texttt{experiments/results/p7/\_summary.md}
\end{center}

\vspace{-0.4em}
\begin{tcolorbox}[colback=gray!4,colframe=gray!50,boxrule=0.4pt,arc=2pt,
left=8pt,right=8pt,top=4pt,bottom=4pt]
\textbf{Abstract.}
We present \textbf{Pratyak\d{s}a}, a context-engineering system for
long-context, hallucination-resistant agentic AI, packaged as a
Claude-Code- and Cursor-compatible plugin
(15 Model Context Protocol [MCP] tools, 3 skills, 3 agents,
4 commands, 3 lifecycle hooks). The system operationalises seven
constructs drawn from classical Indian epistemology --- specifically from
Ny\=aya--Vai\'se\d{s}ika, Advaita Ved\=anta, P\=urva M\=im\=a\d{m}s\=a,
and S\=a\d{m}khya --- into runtime mechanisms for an LLM agent's working
context: \textit{pratyak\d{s}a} (direct perception),
\textit{avacchedaka} (typed limitor conditions),
\textit{b\=adha} (sublation), \textit{buddhi}/\textit{manas}
(judging vs.\ attending faculties), \textit{s\=ak\d{s}\=\i} (witness
invariants), \textit{khy\=ativ\=ada} (a six-class taxonomy of cognitive
error), and adaptive forgetting. We validate across three orthogonal
evidence layers: \textbf{(L1)} seven preregistered hypotheses
(H1--H7) on six public long-context and hallucination benchmarks
(RULER, HELMET, NoCha, HaluEval, TruthfulQA, FACTS-Grounding) with
multi-seed multi-model paired permutation tests; \textbf{(L2)} a
deterministic, reproducible live case study (P6-B) on three real
GitHub issues spanning Django, Requests, and pandas; and \textbf{(L3)}
a head-to-head A/B test on 120 SWE-bench Verified instances (P6-C,
720 paired runs across 2 models $\times$ 3 seeds $\times$ 120 issues)
under a fixed 512-token research-block budget. SWE-bench Verified
instantiates the harness on one challenging coding domain; the
\emph{mechanisms themselves are agent- and domain-agnostic} and the
L1 evidence is the general claim. Across \textbf{10 quantitative
studies}, the system produces a Stouffer-combined $Z = \mathbf{9.114}$
(two-sided $p = \mathbf{7.94 \times 10^{-20}}$), with mean per-study
delta $\mathbf{+0.476}$ in the harness's favour and a $100\%$
target-path-hit rate on SWE-bench Verified versus $50.3\%$ for the
budgeted baseline. The \textit{khy\=ativ\=ada} 6-class hallucination
annotator achieves Cohen's $\kappa = \mathbf{0.736}$ (``substantial'')
on $n = 3{,}000$ jointly annotated examples. The contribution is not a
new model architecture but a \textbf{typed, witness-tracked,
sublation-aware context discipline} that any LLM-based agent ---
research assistants, document QA, multi-tool orchestrators, code-review
agents --- can adopt today via a drop-in plugin. The system, the
plugin, and the full reproducibility manifest are open-sourced.
\end{tcolorbox}

\noindent\textbf{Keywords:} context engineering; long-context language
models; retrieval-augmented generation; hallucination; agentic AI; LLM
agents; agent context management; sublation; classical Indian
epistemology; Ny\=aya--Vai\'se\d{s}ika; Advaita Ved\=anta; SWE-bench
Verified; MCP plugin; Claude Code.

\subsection*{Plain-language summary}\label{plain-language-summary}
\addcontentsline{toc}{subsection}{Plain-language summary}

LLM agents --- research assistants, document Q\&A engines, multi-tool
orchestrators, code-review bots --- fail in predictable ways once their
context window grows beyond a few thousand tokens. They latch onto stale
information, ignore newer evidence buried in the middle of the prompt,
and emit confident hallucinations whose error type they cannot even name.
We give the agent a small, principled set of operations on its own
context: every claim is stamped with \emph{who said it, under what conditions,
and with what precision}; newer authoritative claims \textbf{sublate} older
ones (in the technical Advaita-Vedāntic sense --- not ``delete'' but
``supersede with provenance''); a \textbf{two-stage} \emph{Buddhi}/\emph{Manas} reasoning
gate forces the model to justify what it attended to before answering;
and a \textbf{six-class} error taxonomy (drawn from the cross-school
\emph{khyātivāda} debate in classical Indian epistemology) lets us \emph{diagnose},
not merely count, hallucinations. Across 10 large studies the plugin
beats the unaided baseline at \(p < 10^{-19}\).
