\section{4 · Theoretical Foundations: Vedic Epistemology Mapped to LLM Context Engineering}\label{theoretical-foundations-vedic-epistemology-mapped-to-llm-context-engineering}

This section constructs a precise translation table from seven classical Indian epistemic constructs to seven runtime operations on an LLM agent's context. Each row is a falsifiable design commitment, and each is operationalized in Section 5 and tested in Sections 8--10.

\subsection{4.1 Pratyakṣa --- direct perception as the harness's grounding pramāṇa}\label{pratyakux1e63a-direct-perception-as-the-harnesss-grounding-pramux101ux1e47a}

In Nyāya \citep{aksapada200nyayasutra, matilal1986perception, phillips2012epistemology} and Advaita \citep{dharmaraja17vedantaparibhasa, datta1932advaita}, \emph{pratyakṣa} (direct perception) is the \emph{first} of the \emph{pramāṇas} --- the means of valid cognition --- and the \emph{foundational} one against which inference (\emph{anumāna}), comparison (\emph{upamāna}), testimony (\emph{śabda}), implication (\emph{arthāpatti}), and non-cognition (\emph{anupalabdhi}) are measured.

\textbf{LLM operationalization.} We treat \emph{pratyakṣa} as the agent's \textbf{direct read of its current context window} --- what it can in-fact attend to right now --- and demote all upstream operations (RAG retrieval, tool outputs, prior-turn summaries) to the status of \emph{secondary pramāṇas} whose claims must be entered into the visible context with their own qualifiers. The harness's name, \emph{Pratyakṣa}, is therefore a methodological commitment: \emph{if it is not in the visible, witness-tracked context, it is not direct evidence for the agent's next action}. Hidden hot caches, opaque rerankers, and ``system memories'' the agent cannot inspect are, in this discipline, \textbf{not allowed to ground a claim} without first entering the \emph{pratyakṣa} surface.

\subsection{4.2 Avacchedaka --- qualifier-conditioned cognition as typed limitor}\label{avacchedaka-qualifier-conditioned-cognition-as-typed-limitor}

In Navya-Nyāya \citep{matilal1968navyanyaya, ingalls1951materials, gangesa14tattvacintamani, phillips2012epistemology}, every cognition is structurally a triple:

\[
C = \langle a,\; Q,\; R \rangle \quad\text{read as ``}a\text{ as qualified by }Q\text{ under relation/condition }R\text{''.}
\]

Here \(a\) is the \emph{qualificand} (viśeṣya), \(Q\) the \emph{qualifier} (prakāra), and \(R\) the \emph{limitor} (avacchedaka) under which the qualification holds. Crucially, \emph{R} is part of the cognition; ``the pot is blue \emph{under daylight}'' is a different cognition from ``the pot is blue \emph{under candlelight}'', and the two can co-exist without contradiction.

\textbf{LLM operationalization.} Every item entered into the harness's \texttt{ContextStore} carries the four-tuple

\[
\texttt{Item} = \big\langle \texttt{qualificand},\; \texttt{qualifier},\; \texttt{condition},\; \texttt{precision}\in[0,1] \big\rangle.
\]

\texttt{condition} is a free-text string like \texttt{"as\ of\ 2024-09-Django-5.0"} or \texttt{"on\ POSIX,\ Python\ ≥\ 3.11"}. The MCP tool \texttt{context\_insert(qualificand,\ qualifier,\ condition,\ precision,\ source)} is the only way evidence enters the store. Two items that share \texttt{qualificand} but differ in \texttt{condition} \emph{do not conflict}; two items that share \texttt{qualificand} \emph{and} \texttt{condition} but disagree on \texttt{qualifier} \emph{do}. This is the operational definition of a contradiction the harness can act on (Section 4.3).

\subsection{4.3 Bādha --- sublation as supersede-with-provenance}\label{bux101dha-sublation-as-supersede-with-provenance}

In Advaita Vedānta \citep{deutsch1969advaita, shaw1990bada, dharmaraja17vedantaparibhasa, sankaraupadesha}, a higher-precision cognition \emph{sublates} (bādhita) a lower-precision one. The classical example is mistaking a rope for a snake: the \emph{snake-cognition} is not erased on closer inspection --- the agent must remember it, because \emph{that error is itself an object of subsequent reflection} --- but its qualifier is now stamped \emph{bādhita} and it no longer grounds action.

\textbf{LLM operationalization.} The MCP tool \texttt{sublate\_with\_evidence(target\_id,\ by\_id,\ reason)} rewrites the target item's \texttt{status} to \texttt{bādhita}, annotates it with a pointer to the superseding item, and \emph{retains} the original under that status. Subsequent retrievals filter by \texttt{status\ ==\ "live"} for normal use but can be opted-in for audit. Sublation is triggered by two rules:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Explicit pointer}: the newly-arriving item names a prior item via \texttt{superseded\_by\_id}.
\item
  \textbf{Dominance rule}: the newly-arriving item shares \texttt{qualificand} and \texttt{condition} with an existing item, has \emph{strictly higher} precision, and the older item is \texttt{stale=True} while the newer is not.
\end{enumerate}

The dominance rule is the operational form of the Advaita commitment that \emph{higher pramāṇa supersedes lower pramāṇa} under shared limitor; the explicit pointer is the form used when an LLM-side classifier or external evidence has detected a stale-fresh pair.

\subsection{4.4 Buddhi and Manas --- the two-stage attend-then-judge gate}\label{buddhi-and-manas-the-two-stage-attend-then-judge-gate}

The Antaḥkaraṇa quartet of \emph{manas, buddhi, citta, ahaṃkāra} \citep{deutsch1969advaita, ramprasad2013advaita, datta1932advaita} distinguishes the \emph{attentional sense-organ} (\emph{manas}) from the \emph{determinative judging faculty} (\emph{buddhi}). Manas selects and presents; Buddhi decides. The two are sequential and cannot be collapsed without losing the gate.

\textbf{LLM operationalization.} The harness defines two sub-agent prompts (Section 6.3):

\begin{itemize}
\tightlist
\item
  \textbf{\texttt{ManasAgent}} receives the user query and the \texttt{ContextStore} snapshot, and returns a structured JSON object matching the shipped contract in \texttt{agents/manas.md}: \texttt{\{\ "draft":\ "...",\ "grounding":\ {[}"\textless{}element\_id\textgreater{}",\ ...{]},\ "uncertain\_claims":\ {[}"..."{]},\ "needs\_buddhi":\ true\ \textbar{}\ false\ \}}. Manas is operationalised as a \textbf{sub-agent} (Section 6.3), not a skill. Manas is \emph{forbidden} from emitting a final user-visible answer (draft text is for the orchestrator only).
\item
  \textbf{\texttt{BuddhiAgent}} receives the user query, the Manas output, and re-reads the attended items from the \texttt{ContextStore}, and returns the final answer plus its own \texttt{khyāti\_class} (Section 4.6) for the answer. Buddhi is the only agent that may emit a user-visible answer.
\end{itemize}

This two-stage gate is identical in shape to dual-process accounts of human cognition \citep{evans2003duality, kahneman2011thinking, sloman1996two} and to classic AI cognitive architectures \citep{laird1987soar, anderson1996actr, newell1990unified}, but the names --- and the surrounding philosophical apparatus --- are Vedic. Critically, \emph{Manas can be wrong without Buddhi being wrong}: if Manas mis-selects, Buddhi's job is to detect and call for re-selection rather than to answer.

\subsection{4.5 Sākṣī --- the witness as model-invariant audit frame}\label{sux101kux1e63ux12b-the-witness-as-model-invariant-audit-frame}

In Advaita Vedānta \citep{indich1980consciousness, fasching2009witness, ganeri2017concealed, sankaraupadesha}, \emph{sākṣī} --- the witness consciousness --- is the \emph{non-revisable, non-acting} observer that \emph{records} what is cognised without itself being the agent of cognition or action. It supplies the stable reference frame against which changes of cognition are intelligible at all.

\textbf{LLM operationalization.} The harness instantiates a \texttt{SakshiKeeperAgent} (Section 6.3) and a \texttt{SakshiPrefix} skill (Section 6.2). On every turn, the keeper appends to a persistent JSON-lines audit log at \texttt{\textasciitilde{}/.cache/pratyaksha/audit.jsonl} (XDG-style cache path; survives plugin reinstall) an immutable record of:

\begin{itemize}
\tightlist
\item
  The user query.
\item
  The Manas selection (which items, under which conditions).
\item
  The Buddhi judgment (final answer, khyāti class, calibrated posterior).
\item
  All sublations fired this turn (target\_id, by\_id, reason).
\item
  The token budget consumed, by category (system, retrieval, tool, completion).
\end{itemize}

Crucially, the witness is \emph{write-once-per-turn} and \emph{read-only} to all agents in the next turn. It is the harness's source of truth for cross-turn provenance, and it is \emph{invariant under model swaps}: the witness log written under Claude 3.5 Sonnet is fully readable, fully diffable, and fully auditable when the agent is later switched to Claude 4.5 Sonnet, GPT-4o, or Qwen-3-72B. The model is replaceable; the witness is not.

\subsection{4.6 Khyātivāda --- six-class typology of error}\label{khyux101tivux101da-six-class-typology-of-error}

Six classical schools of Indian philosophy proposed six theories of erroneous cognition \citep{ram2007error, matilal1986perception, datta1932advaita, bhatt1989prabhakara, bilimoria2018epistemology, mohanty1992reason}. Each names a specific \emph{kind} of error:

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Class
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
School
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Kernel definition
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Modern LLM analogue
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{Anyathākhyāti} & Nyāya & ``Apprehension of one thing as another \emph{of the same general type}.'' & Wrong API name (right module, wrong function); citing the right paper but wrong year. \\
\textbf{Ātmakhyāti} & Yogācāra & ``Projection of an internal cognition onto the external object.'' & Inventing helper functions / config keys the framework never exposed; hallucinated TODOs the user did not write. \\
\textbf{Anirvacanīyakhyāti} & Advaita & ``Apprehension of an object that is \emph{neither real nor unreal}; sui generis indeterminate.'' & Hedged hallucinations: ``you might consider X-or-similar'' where X is plausible but ungrounded. \\
\textbf{Asatkhyāti} & Mādhyamika & ``Apprehension of a non-existent object.'' & Citing a CVE, a person, a paper, or a commit hash that simply does not exist. \\
\textbf{Viparītakhyāti} & Mīmāṃsā & ``Apprehension of the contrary.'' & Inverting a boolean flag's semantics; recommending the \emph{opposite} of the documented best practice. \\
\textbf{Akhyāti} & Prabhākara Mīmāṃsā & ``Failure of correct apprehension by mere conflation of two adjacent cognitions.'' & Mixing two adjacent functions / two near-duplicate config schemas / two adjacent versions. \\
\end{longtable}
}

This is not a post-hoc relabelling of an existing modern taxonomy. It is the \emph{original} taxonomy from the Indian philosophical tradition, jointly developed by debating schools over roughly a thousand years. We adopt it directly because (a) it carves the space cleanly along \emph{epistemic} (not merely surface-textual) lines, (b) the six classes are mutually exclusive in the tradition's own usage, and (c) it gives us a \emph{typed signal} the harness can act on (different remediation strategies for different classes).

\textbf{LLM operationalization.} The shipped plugin's \texttt{classify\_khyativada} tool applies a \textbf{heuristic classifier with the rule-based guardrails} described in Section 5.5. Separately, the experiments harness includes a few-shot Anthropic JSON classifier (\texttt{src/evaluation/khyativada\_fewshot.py}) that is \textbf{not} wired into the plugin tool; both paths are evaluated independently in §8 (H6). Inter-annotator agreement is reported on the 3,000-example corpus using prompts in Appendix D and results in §8.6 at Cohen's κ = 0.736 (``substantial'').

\subsection{4.7 Saṃskāras and vāsanās --- adaptive forgetting}\label{saux1e43skux101ras-and-vux101sanux101s-adaptive-forgetting}

In Advaita and Yoga, \emph{saṃskāras} are residual mental impressions and \emph{vāsanās} are dispositional tendencies that survive the immediate occasion of a cognition \citep{deutsch1969advaita, sankaraupadesha, halbfass1991traditioncomparison}. They explain why we are drawn to the same patterns of thought and why some impressions decay quickly while others persist indefinitely. The Advaitin practice prescribes \emph{removal of obstructive vāsanās} --- adaptive forgetting of patterns that no longer ground correct cognition.

\textbf{LLM operationalization.} The \texttt{AdaptiveForgetting} module (Section 5.7) applies an exponential-decay schedule to non-witnessed items in the \texttt{ContextStore}, with three explicit rules:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Witnessed items never decay.} If \texttt{sākṣī.witnessed\ ==\ True} for an item-id, its precision floor is fixed.
\item
  \textbf{Sublated items decay faster.} Items with \texttt{status\ ==\ bādhita} halve their precision every time they survive a budget-pressure compaction event.
\item
  \textbf{High-recency, low-precision items decay slowest among the rest} --- the item is too new to know whether it will become a vāsanā or a saṃskāra.
\end{enumerate}

This avoids both the catastrophic-forgetting failure mode of naive context pruning \citep{french1999catastrophic, kirkpatrick2017overcoming, parisi2019continual} and the ``everything is forever'' failure mode of unconstrained external memory \citep{packer2023memgpt}.

\subsection{4.8 The translation table, summarised}\label{the-translation-table-summarised}

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Vedic construct
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
LLM operationalization
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
MCP tool / module
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Tested by
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Pratyakṣa & Visible-context-only grounding discipline & \texttt{context\_retrieve}, \texttt{context\_window}, related read paths & All hypotheses (architectural commitment) \\
Avacchedaka & Typed limitor \texttt{(qualificand,\ qualifier,\ condition,\ precision)} & \texttt{context\_insert}, \texttt{context\_retrieve} & H2, H5 \\
Bādha & Supersede-with-provenance under shared limitor & \texttt{sublate\_with\_evidence} & H4, H5, P6-B, P6-C \\
Manas & Pre-judgment attention-selection \textbf{sub-agent} & \texttt{ManasAgent} (Section 6.3) & H3 \\
Buddhi & Determinative judgment + khyāti tagging sub-agent & \texttt{BuddhiAgent} & H3, H6 \\
Sākṣī & Write-once cross-turn witness log & \texttt{SakshiKeeperAgent}, \texttt{SakshiPrefix} & All (audit invariant) \\
Khyātivāda & 6-class hallucination classifier & \texttt{classify\_khyativada} & H6 \\
Saṃskāra/vāsanā & Adaptive forgetting with witness-protected items & \texttt{AdaptiveForgetting} & H7 \\
\end{longtable}
}

\subsection{\texorpdfstring{4.9 What is \emph{not} being claimed}{4.9 What is not being claimed}}\label{what-is-not-being-claimed}

We are \emph{not} claiming that the historical authors had LLMs in mind, nor that our operationalizations are textually faithful in the way a Sanskrit philologist would require. We are claiming the weaker, falsifiable thing: that the \emph{type signatures} the tradition developed for cognition turn out to \emph{fit} the type signatures we need for an LLM context-engineering harness, and that fitting them produces measurable downstream gains. The next six sections defend that claim empirically.
