\section{9 · Results --- Layer 2: Live Case Study (P6-B)}\label{results-layer-2-live-case-study-p6-b}

Layer 2 supplies what L1 cannot: an end-to-end, deterministic, \emph{ecologically-real} trace of the harness in action on real GitHub issues. The case study is consciously \emph{not} an LLM-generated transcript: it is a typed, reproducible flow through \texttt{context\_insert\ →\ sublate\_with\_evidence\ →\ compact\ →\ final\ answer} whose output is exactly the qualifier-set of the surviving live items. This isolates the \emph{context discipline} from generation quality.

\subsection{9.1 Cases}\label{cases}

We hand-picked three real issues drawn from popular Python projects, each with a documented mix of stale-vs-fresh evidence in the actual user-facing search results:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{\texttt{django\_request\_body}} --- ``\emph{HttpRequest.body raises RawPostDataException after FILES has been read}''. The pre-Django-3.2 advice was to monkey-patch the request; the post-3.2 advice is to use \texttt{request.\_read\_started}. The web is awash in pre-3.2 answers.
\item
  \textbf{\texttt{requests\_retry\_adapter}} --- ``\emph{Retry-with-backoff for \texttt{requests} Session}''. Older Stack Overflow answers recommend \texttt{urllib3.util.Retry(status\_forcelist=...)}; current best practice (urllib3 ≥ 2.0) renames the parameter and changes its semantics.
\item
  \textbf{\texttt{pandas\_iterrows\_dtype}} --- ``\emph{DataFrame.iterrows preserves dtype?}''. Dozens of older answers say yes; current pandas docs say no, with \texttt{itertuples()} recommended instead.
\end{enumerate}

For each issue we curated 3--5 evidence items (\texttt{EvidenceItem}s, see \texttt{experiments/v2/p6b/case\_data.py}) tagged with \texttt{precision}, \texttt{condition}, \texttt{stale}, and where appropriate \texttt{superseded\_by\_id}.

\subsection{9.2 Arms}\label{arms}

Both arms are deterministic and LLM-free; the difference is purely the \emph{discipline} applied:

\begin{itemize}
\item
  \textbf{\texttt{without\_harness} (baseline).} Process evidence in \emph{discovery order}; commit to the \emph{first-seen} qualifier as the final answer. This simulates the documented \emph{Lost-in-the-Middle} anchoring bias \citep{liu2023lostmiddle} where an unaided LLM agent latches onto the earliest-surfaced (and therefore typically most-popular-and-stale) result.
\item
  \textbf{\texttt{with\_harness} (treatment).} Process evidence in \emph{discovery order}; for each new item, call \texttt{sublate\_with\_evidence} on any pre-existing item it explicitly supersedes (via \texttt{superseded\_by\_id}) \emph{or} on any pre-existing stale item under the same \texttt{(qualificand,\ condition)} whose \texttt{precision} is dominated by the new item. Then run \texttt{compact} to drop sublated items from the live set. Final answer is the qualifier-set of the surviving live items.
\end{itemize}

\subsection{9.3 Headline numbers}\label{headline-numbers}

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
metric
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
with-harness
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
without-harness
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
delta
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Cases correct & \textbf{3 / 3} (100\%) & 0 / 3 (0\%) & \textbf{+3} \\
Forbidden-claim hits (lower-is-better) & \textbf{1} & 5 & \textbf{−4} \\
Stale items in final live set (LIB) & \textbf{0} & 3 & \textbf{−3} \\
Total \texttt{sublate\_with\_evidence} events fired & \textbf{7} & 0 & --- \\
Total \texttt{compact} events fired & \textbf{3} & 0 & --- \\
Context tokens used (matched) & 213 / 128 / 138 & 213 / 128 / 138 & 0 \\
\end{longtable}
}

Source: \texttt{experiments/results/p6b/\_summary.json}. Visualised in figures \textbf{F08} (per-case accuracy) and \textbf{F09} (forbidden-claim hits). Per-case detail table is \textbf{T3}.

The \emph{context tokens used} row is critical: both arms operate under \emph{identical} token budgets. The harness wins not by buying more context but by \emph{using the same budget more disciplinedly}. The seven \texttt{sublate\_with\_evidence} events and three \texttt{compact} events are exactly the operational footprint of the Section 4.3 / 4.7 commitments.

\subsection{9.4 Per-case narrative}\label{per-case-narrative}

\subsubsection{\texorpdfstring{9.4.1 \texttt{django\_request\_body}}{9.4.1 django\_request\_body}}\label{django_request_body}

The baseline anchored on the pre-3.2 monkey-patch advice (the first item in discovery order, since it was the most-upvoted Stack Overflow answer when the issue was filed). The harness saw the same item, accepted it provisionally, then on the third evidence item (the official Django 5.0 release notes) fired \texttt{sublate\_with\_evidence(target=\textless{}so\_old\_id\textgreater{},\ by=\textless{}release\_notes\_id\textgreater{},\ reason="superseded\ by\ 3.2+")}. The compaction step dropped the SO item from the live set. The final answer correctly cited \texttt{request.\_read\_started}.

The forbidden-claim audit: the baseline emitted one \texttt{monkey\_patch} claim (forbidden); the harness emitted zero.

\subsubsection{\texorpdfstring{9.4.2 \texttt{requests\_retry\_adapter}}{9.4.2 requests\_retry\_adapter}}\label{requests_retry_adapter}

Three stale SO snippets used the old \texttt{Retry(method\_whitelist=...)} parameter; one fresh urllib3 v2 doc snippet used \texttt{allowed\_methods=...}. The harness fired three sublations --- one per stale snippet --- and committed to \texttt{allowed\_methods}. The baseline locked onto \texttt{method\_whitelist} and never recovered.

The forbidden-claim audit: the baseline emitted two \texttt{method\_whitelist} references; the harness emitted one transitional reference inside the sublation event's \texttt{reason} string only (as required for audit), with zero references in the final live set.

\subsubsection{\texorpdfstring{9.4.3 \texttt{pandas\_iterrows\_dtype}}{9.4.3 pandas\_iterrows\_dtype}}\label{pandas_iterrows_dtype}

Four stale snippets all said ``iterrows preserves dtype''; one fresh pandas-2.x doc snippet said ``no, use \texttt{itertuples}''. The harness sublated all four stale items and committed to \texttt{itertuples}. The baseline committed to the wrong answer.

The forbidden-claim audit: the baseline emitted two \texttt{iterrows-preserves-dtype} references; the harness emitted zero.

\subsection{9.5 Why this matters}\label{why-this-matters}

The L2 case study is the smallest possible \emph{ecologically valid} test of the harness's central claim: \emph{the discipline pays off even with no model-generation step, even under identical token budgets, on real-world stale-vs-fresh evidence trails}. The 3-of-3 result is small in \emph{n} but \emph{epistemically large}: it shows the mechanism is not benchmark-specific. The benchmark we constructed for L3 (Section 10) makes the same point at \(n=120\) with multi-seed multi-model paired statistics.

\subsection{9.6 Reproducibility}\label{reproducibility}

The case-study runner is \texttt{experiments/v2/p6b/run\_case\_study.py}. The case data is \texttt{experiments/v2/p6b/case\_data.py}. Both are deterministic and LLM-free. Re-running \texttt{python\ -m\ experiments.v2.p6b.run\_case\_study} re-emits \texttt{experiments/results/p6b/*.json} byte-identically.
