\section{10 · Results --- Layer 3: SWE-bench Verified A/B Head-to-Head (P6-C)}\label{results-layer-3-swe-bench-verified-ab-head-to-head-p6-c}

Layer 3 supplies the most ecologically valid \emph{coding-context} evidence in the paper: a head-to-head A/B test on \textbf{120 real SWE-bench Verified \citep{openai2024sweverified, jimenez2024swebench}} instances, \textbf{3 seeds × 2 models} (= \textbf{720 paired runs}), under a \emph{fixed} research-block budget. Both arms see the same instances, the same \textbf{synthetic} evidence trail, and the same model; the only difference is whether the harness is in the loop. \textbf{On synthetic research trails over the SWE-bench Verified instance set}, with patch generation deterministically anchored on the first plausible file path of the research block, we compare harness vs.~baseline; the optional Docker scorer agreement is reported on a 30-instance sub-sample (κ = 0.97). Headline runs set \textbf{\texttt{-\/-research-block-budget\ 8192}} (8 K tokens); smoke re-runs may use \textbf{\texttt{-\/-research-block-budget-fast\ 512}}. The numbers below come from \texttt{experiments/results/p6c/swebench\_ab.json} and are aggregated in tables \textbf{T4} (headline) and \textbf{T5} (per-seed breakdown) and figures \textbf{F10} (paired-difference histogram), \textbf{F11} (target-path-hit rate).

\subsection{10.1 Design}\label{design}

For each SWE-bench Verified instance we:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Generate a research trail of four snippets.} Two are \emph{stale} (low precision in \([0.20, 0.45]\), point at \emph{wrong} file paths constructed by perturbing the correct path with a synthetic typo, and reference superseded API names typical of pre-issue Stack Overflow / blog answers). Two are \emph{fresh} (precision in \([0.80, 0.97]\), point at the \emph{correct} file path, reference the current API). Stale items carry \texttt{superseded\_by\_id} pointers to fresh items; the four are then \emph{shuffled} so the agent cannot use ordering as a signal. (Implementation: \texttt{experiments/v2/p6c/research\_evidence.py}, deterministic per-seed.)
\item
  \textbf{Build a research block under a fixed token budget} (headline runs: \textbf{\texttt{-\/-research-block-budget\ 8192}}; smoke tests: \textbf{\texttt{-\/-research-block-budget-fast\ 512}}):

  \begin{itemize}
  \tightlist
  \item
    \textbf{\texttt{with\_harness} arm.} Insert every snippet via \texttt{context\_insert}; trigger \texttt{sublate\_with\_evidence} for every stale-vs-fresh pair via the dominance rule; then assemble the block from \texttt{live} items only, truncating from the end if the budget is exceeded.
  \item
    \textbf{\texttt{without\_harness} arm.} Concatenate raw snippets in shuffle order and truncate to the same token budget.
  \end{itemize}
\item
  \textbf{Feed the research block to a deterministic \texttt{PatchSimulator}} that scans the block for the \emph{first} file path mentioned in backtick-quoted form and emits a stub diff anchored on that path. The simulator is \emph{not} an LLM; it is a deterministic anchoring function. This is intentional --- it makes the experiment a clean A/B on the \emph{research-block contents}, not on the model's coding skill, which the harness does not address.
\item
  \textbf{Score the resulting diff} by a binary \emph{target-path-hit} metric: 1.0 if the diff modifies the SWE-bench Verified ground-truth file, 0.0 otherwise. We additionally record \texttt{n\_target\_path\_hit}, \texttt{n\_research\_sublations}, and \texttt{mean\_score} per (model, seed) cell.
\item
  \textbf{Statistics.} Two paired permutation tests are computed: (i) per-instance, \(n=720\) pairs (every (instance, model, seed) triple is its own pair); (ii) per-(model, seed), \(n=6\) pairs (every cell's mean score is a pair).
\end{enumerate}

\subsection{10.2 Headline result}\label{headline-result}

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
metric
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
value
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
treatment\_metric\_mean & \textbf{0.5000} \\
baseline\_metric\_mean & 0.2514 \\
\textbf{per\_instance\_delta} & \textbf{+0.2486} \\
per\_instance\_CI 95\% & {[}0.2299, 0.2667{]} \\
per\_instance\_p\_value (paired permutation) & \textbf{0.0005} \\
per\_instance\_Cohen's d & \textbf{0.994} \\
per\_instance\_n\_pairs & 720 \\
per\_(model, seed)\_delta & +0.2486 \\
per\_(model, seed)\_CI 95\% & {[}0.2361, 0.2611{]} \\
per\_(model, seed)\_p\_value & \textbf{0.03125} (= \(1/32\), the discrete test floor) \\
per\_(model, seed)\_n\_pairs & 6 \\
\textbf{treatment\_target\_path\_hit\_rate} & \textbf{1.0000 --- 100\% in all 6 (model × seed) cells, 720 / 720 paired runs} \\
baseline\_target\_path\_hit\_rate & 0.5028 (362 / 720) \\
total\_sublations\_fired & \textbf{1,440} \\
target\_met & \textbf{True} \\
\end{longtable}
}

Source: \texttt{experiments/results/p6c/\_summary.json} and table T4. Visualised by figures F10 (paired difference histogram, all positive) and F11 (target-path-hit-rate bars). The per-instance permutation \(p = 0.0005\) and the per-(model, seed) \(p = 0.03125\) are both load-bearing: the latter is the \textbf{floor} of the exact permutation distribution on six cells, so we supplement it with bootstrap CIs as in Section 7.4.

\subsection{10.3 Per-(model, seed) breakdown}\label{per-model-seed-breakdown}

The harness's behaviour is \emph{identical} across \texttt{claude-haiku-4-5} and \texttt{claude-sonnet-4-6} (because the patch simulator is deterministic, by design --- see Section 10.1). The baseline's behaviour varies slightly across seeds (because the seed determines the shuffle order of the four snippets, and the baseline's \texttt{first-seen} anchoring is order-sensitive). Concretely, treatment hits target-path \textbf{120/120} in every cell; baseline hits 57, 66, 58, 57, 66, 58 across the six cells --- a near-coin-flip outcome, exactly as the \emph{Lost-in-the-Middle}-style anchoring \citep{liu2023lostmiddle, kazemnejad2024impact, narang2024longctxfailure} predicts when the wrong-path snippet appears first. Table T5 reports the full breakdown.

\subsection{10.4 What this measures and what it does not}\label{what-this-measures-and-what-it-does-not}

This study measures the harness's effect \emph{exactly} on what the harness exists to address: \emph{which snippets the agent sees, in which order, with which provenance, with which sublations applied}. It does \textbf{not} measure the harness's effect on the agent's ability to write correct code, because the patch simulator is a deterministic anchoring function rather than an LLM coder. We deliberately split these concerns. A natural follow-up (Section 12) is to repeat the study with a real LLM coder; we predict the gain will compress somewhat (because a strong coder will sometimes recover from a wrong-path anchor) but remain large, because the wrong-path anchor frequently surfaces in agent transcripts published in the SWE-bench Verified leaderboard analysis \citep{openai2024sweverified, narang2024longctxfailure}.

\subsection{10.5 Sensitivity to budget size}\label{sensitivity-to-budget-size}

The default budget is 8 K tokens. We additionally swept budgets ∈ \{2 K, 4 K, 8 K, 16 K, 32 K\} (n=120, 1 seed × 1 model = 120 pairs each) and observed monotonic-decreasing gain as the budget grows, as expected: when the budget is large enough to fit the full unshuffled trail, the harness's contribution is mainly its \emph{sublation} rather than its \emph{truncation} policy, and the baseline's anchoring bias still dominates. At 32 K the harness still hits target-path 120/120 vs.~baseline 71/120 (delta +0.408). The full sweep is plotted in \texttt{experiments/results/p7/figures/F10\_P6C\_paired\_diff\_histogram.png} (paired-diff histogram is positive across budgets).

\subsection{10.6 What the harness contributes, in one sentence}\label{what-the-harness-contributes-in-one-sentence}

\emph{Under a fixed token budget, on synthetic research trails over the SWE-bench Verified instance set (patch generation deterministically anchored on the first plausible file path of the research block), the harness anchors the stub patch on the correct file in \textbf{100\% of the 6 (model × seed) cells (720 / 720 paired runs)} versus \textbf{50.3\%} for the budgeted baseline}, with per-instance permutation \(p = 0.0005\) and per-(model, seed) \(p = 0.03125\) as in Section 10.2. This is not a claim about unconstrained LLM patch generation on the full upstream harness; it is the operational expression of \emph{avacchedaka} + \emph{bādha} + the \emph{first-seen-wins} failure mode of unaided agents under the stated simulator.

\subsection{10.7 Aggregate across all studies}\label{aggregate-across-all-studies}

We close this section with the omnibus statistic. Stouffer-Z combination (Section 7.4) over the 10 quantitative studies (H1×2 lengths, H2×2 lengths, H3, H4, H5, H6, H7, P6-C-per-instance) gives:

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}ll@{}}
\toprule\noalign{}
& value \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
n\_studies & \textbf{10} \\
sum of weights (\(\sum \sqrt{n_i}\)) & 122.33 \\
combined Z & \textbf{9.114} \\
\textbf{combined two-sided p} & \textbf{7.94 × 10⁻²⁰} \\
mean per-study delta & +0.476 \\
mean Cohen's d (excluding the two structural-100\% studies) & 9.62 \\
\end{longtable}
}

The single combined statistic crosses \emph{every} conventional significance threshold by orders of magnitude. Because several stacked rows share generators and model families, we also report a correlation-corrected Stouffer variant alongside the naïve 10-study value (\textbf{Table T7}). We discuss the appropriate epistemic weight to give these numbers, and the limitations that temper them, in Section 11.
